### 2023-03-16 outputs/paper/paper.qmd
set-up

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: LINK."
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
```

# Introduction

You can and should cross-reference sections and sub-sections. For instance, @sec-data and @sec-first-point.

# Data {#sec-data}

Our data is of penguins (@fig-bills).

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

model<-lm(educ~mapres10+papres10+paeduc+maeduc,data=raw_data)
summary(model)
pairs(raw_data)
# ggplot(penguins, aes(x = island, fill = species)) +
#   geom_bar(alpha = 0.8) +
#   scale_fill_manual(values = c("darkorange","purple","cyan4"),
#                     guide = "none") +
#   theme_minimal() +
#   facet_wrap(~species, ncol = 1) +
#   coord_flip()
```

Talk more about it.

Also bills and their average (@fig-billssssss). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work.)

```{r}
#| label: fig-billssssss
#| fig-cap: More bills of penguins
#| echo: false
#| fig-width: 8
#| fig-height: 4

ggplot(raw_data, aes(x = papres10, y = educ)) +
  geom_point(alpha=0.3)
```

Talk way more about it.

# Model

$$
Pr(\theta | y) = \frac{Pr(y | \theta) Pr(\theta)}{Pr(y)}
$$ {#eq-bayes}

@eq-bayes seems useful, eh?

Here's a dumb example of how to use some references: In paper we run our analysis in `R` [@citeR]. We also use the `tidyverse` which was written by @thereferencecanbewhatever If we were interested in baseball data then @citeLahman could be useful.

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional details

\newpage

# References

### 2023-03-16 scripts/02-data_cleaning.R

@@ -1,0 +1,45 @@
+#### Preamble ####
+# Purpose: Cleans.... [...UPDATE THIS...]
+# Author: Rohan Alexander [...UPDATE THIS...]
+# Data: 11 February 2023 [...UPDATE THIS...]
+# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# License: MIT
+# Pre-requisites: [...UPDATE THIS...]
+# Any other information needed? [...UPDATE THIS...]
+
+
+#### Workspace setup ####
+library(haven)
+library(tidyverse)
+library(dplyr)
+
+#### Clean data ####
+raw_data <- read_dta("inputs/data/GSS2021.dta")
+
+raw_data <-
+  labelled::to_factor(raw_data)
+
+raw_data <- 
+  raw_data %>% 
+  select(mapres10,papres10,paeduc,maeduc,educ)
+
+#change class
+raw_data$mapres10 <- as.character(raw_data$mapres10)
+raw_data$papres10 <- as.character(raw_data$papres10)
+raw_data$maeduc <- as.character(raw_data$maeduc)
+raw_data$paeduc <- as.character(raw_data$paeduc)
+raw_data$educ <- as.character(raw_data$educ)
+
+#change no formal to 0
+raw_data<-raw_data%>%mutate_all(~ifelse(. == "no formal schooling", 0, .))%>%
+  mutate_all(as.numeric)
+
+#group variables into levels
+
+
+
+#### Save data ####
+# [...UPDATE THIS...]
+# change cleaned_data to whatever name you end up with at the end of cleaning
+write_csv(cleaned_data, "outputs/data/cleaned_data.csv")
+

### 2023-03-16 outputs/paper/paper.qmd
Add level of prestige in the data cleaning add plots

@@
 date: "`r Sys.time()`"
-date-format: "D MMMM YYYY"
+date-format: "12-03-2023"
 abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
@@
 #| message: false
-
+library(broom)
+library(knitr)
+library(kableExtra)
 library(tidyverse)
+library(dplyr)
 ```
@@
 
-model<-lm(educ~mapres10+papres10+paeduc+maeduc,data=raw_data)
-summary(model)
-pairs(raw_data)
-# ggplot(penguins, aes(x = island, fill = species)) +
-#   geom_bar(alpha = 0.8) +
-#   scale_fill_manual(values = c("darkorange","purple","cyan4"),
-#                     guide = "none") +
-#   theme_minimal() +
-#   facet_wrap(~species, ncol = 1) +
-#   coord_flip()
+# Fit the model
+model <- lm(educ ~ mapres10 + papres10 + paeduc + maeduc, data = cleaned_data)
+
+# Get the summary statistics and tidy them up
+tidy_summary <- tidy(summary(model))
+
+# Create a data frame for the tidy summary
+summary_df <- as.data.frame(tidy_summary)
+
+# # Rename the row names
+# row_names <- c("Intercept", "Mother's Prestiges Score", "Father's Prestiges Score", "Father's Education", "Mother's Education")
+# rownames(summary_df) <- row_names
+
+# Rename the column names
+col_names <- c("Term", "Estimate", "Std. Error", "t value", "Pr(>|t|)")
+colnames(summary_df) <- col_names
+
+# Print the summary table with kableExtra
+library(kableExtra)
+kable(summary_df, align = c("l", "r", "r", "r", "r"), 
+      caption = "Summary of Linear Regression Model",) %>%
+  kable_styling(bootstrap_options = "striped", full_width = FALSE)
+
+```
+
+```{r}
+#| label: fig-billssssss
+#| fig-cap: More bills of penguins
+#| echo: false
+#| fig-width: 8
+#| fig-height: 4
+ggplot(data = cleaned_data, aes(x = educ)) +
+
+  geom_histogram(aes(y = ..density..), binwidth = 2, fill = "white") +
+  geom_density(alpha = .2, fill = "lightblue") +
+
+  # geom_histogram(bins = 15, fill = "lightblue", color = "white") +
+  labs(title = "Histogram of Respondents' Education", x = "Education (in Years)", y = "Desensity") +
+   scale_x_continuous(breaks = seq(0, 20, by = 1))+
+  theme_bw()
+```
+
+```{r}
+#| label: fig-billssssss
+#| fig-cap: More bills of penguins
+#| echo: false
+#| fig-width: 8
+#| fig-height: 4
+ggplot(cleaned_data, aes(x = mapres10)) +
+  geom_histogram(aes(y = ..density.., fill = "mapres10"), alpha = 0.8, binwidth = 2,color = "black") +
+  geom_histogram(aes(x = papres10, y = ..density.., fill = "paeduc"), alpha = 0.4, binwidth = 2,color = "black") +
+  scale_fill_manual(values = c("mapres10" = "lightblue", "paeduc" = "red"), name = "Variables") +
+  labs(title = "Difference in Mother's and Father's Prestige Score",
+       x = "Prestige Score",
+       y = "Density")+
+  theme_bw()
+```
+
+```{r}
+#| label: fig-billssssss
+#| fig-cap: More bills of penguins
+#| echo: false
+#| fig-width: 8
+#| fig-height: 4
+ggplot(cleaned_data, aes(x = maeduc)) +
+  geom_histogram(aes(y = ..density.., fill = "maeduc"), alpha = 0.8, binwidth = 1, color = "black") +
+  geom_histogram(aes(x = paeduc, y = ..density.., fill = "paeduc"), alpha = 0.4, binwidth = 1, color = "black") +
+  scale_fill_manual(values = c("maeduc" = "lightblue", "paeduc" = "red"), name = "Variables") +
+  labs(title = "Differene in Mother's Education and Father's Education",
+       x = "Education (in Years)",
+       y = "Density") +
+  theme_bw()
 ```
@@
 
-ggplot(raw_data, aes(x = papres10, y = educ)) +
-  geom_point(alpha=0.3)
+ggplot(cleaned_data, aes(x = educ, fill = factor(father_factor))) +
+  geom_histogram(position = "dodge", binwidth = 1.5, alpha = 0.5, color = "black")+
+   labs(title = "Father's Prestige Score and Education",
+       x = "Education (in Years)",
+       y = "Count",
+       fill = "Level of Prestige Score")+
+  theme_bw()
+```
+
+```{r}
+#| label: fig-billssssss
+#| fig-cap: More bills of penguins
+#| echo: false
+#| fig-width: 8
+#| fig-height: 4
+ggplot(cleaned_data, aes(x = educ, fill = factor(mother_factor))) +
+  geom_histogram(position = "dodge", binwidth = 1.5, alpha = 0.5, color = "black")+
+     labs(title = "Mother's Prestige Score and Education",
+       x = "Education (in Years)",
+       y = "Count",
+       fill = "Level of Prestige Score")+
+  theme_bw()
 ```
@@
 
-\appendix
-
-# Appendix {-}
-
-# Additional details
-
-\newpage
-
 # References

### 2023-03-16 scripts/02-data_cleaning.R

@@ -1,11 +1,11 @@
 #### Preamble ####
-# Purpose: Cleans.... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Data: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Data cleaning for GSS 2021
+# Author: Ruibo Sun
+# Data: 16 March 2023
+# Contact: ruibo.sun@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: Download all the packges under the Workspace section
+# Any other information needed? None
 
 
 #### Workspace setup ####
@@ -31,15 +31,29 @@ raw_data$paeduc <- as.character(raw_data$paeduc)
 raw_data$educ <- as.character(raw_data$educ)
 
 #change no formal to 0
-raw_data<-raw_data%>%mutate_all(~ifelse(. == "no formal schooling", 0, .))%>%
+cleaned_data<-raw_data%>%mutate_all(~ifelse(. == "no formal schooling", 0, .))%>%
   mutate_all(as.numeric)
 
 #group variables into levels
 
+mepa<-median(cleaned_data$papres10,na.rm=TRUE)
+mema<-median(cleaned_data$mapres10,na.rm=TRUE)
+
+cleaned_data <- cleaned_data %>%
+  mutate(father_factor = case_when(
+    papres10 < mepa ~ "Lower than Median",
+    papres10 >= mepa ~ "Above and Equal to Median"
+  ))
+
+cleaned_data <- cleaned_data %>%
+  mutate(mother_factor = case_when(
+    mapres10 < mema ~ "Lower than Median",
+    mapres10 >= mema ~ "Above and Equal to Median"
+  ))
 
 
 #### Save data ####
 # [...UPDATE THIS...]
 # change cleaned_data to whatever name you end up with at the end of cleaning
-write_csv(cleaned_data, "outputs/data/cleaned_data.csv")
+write_csv(cleaned_data, "inputs/data/GSS2021.dta")
 

### 2023-03-17 outputs/paper/paper.qmd
select more variables change the plots and tables

--- 2023-03-16 outputs/paper/paper.qmd
@@
 author: 
-  - First author
-  - Another author
-thanks: "Code and data are available at: LINK."
+  - Ruibo Sun
+thanks: "https://github.com/ruibosun/how-parents-affect-childrens-education"
 date: "`r Sys.time()`"
-date-format: "D MMMM YYYY"
+date-format: "12-03-2023"
 abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
@@
 number-sections: true
-bibliography: references.bib
+#bibliography: references.bib
 ---
@@
 #| message: false
+library(broom)
+library(knitr)
+library(kableExtra)
+library(tidyverse)
+library(dplyr)
+library(here)
+library(haven)
+library(scales)
+library(dplyr)
+library(purrr)
+library(forcats)
+```
 
-library(tidyverse)
+```{r,echo=FALSE}
+cleaned_data <- read_csv(
+  file = here("inputs/data/GSS2021 2.csv"))
 ```
@@
 ```{r}
-#| label: fig-bills
+#| label: fig-bills1
 #| fig-cap: Bills of penguins
@@
 
-model<-lm(educ~mapres10+papres10+paeduc+maeduc,data=raw_data)
-summary(model)
-pairs(raw_data)
-# ggplot(penguins, aes(x = island, fill = species)) +
-#   geom_bar(alpha = 0.8) +
-#   scale_fill_manual(values = c("darkorange","purple","cyan4"),
-#                     guide = "none") +
-#   theme_minimal() +
-#   facet_wrap(~species, ncol = 1) +
-#   coord_flip()
+# Fit the model
+model <- lm(educ ~ mapres10 + papres10 + paeduc + maeduc, data = cleaned_data)
+
+# Get the summary statistics and tidy them up
+tidy_summary <- tidy(summary(model))
+
+# Create a data frame for the tidy summary
+summary_df <- as.data.frame(tidy_summary)
+
+# # Rename the row names
+# row_names <- c("Intercept", "Mother's Prestiges Score", "Father's Prestiges Score", "Father's Education", "Mother's Education")
+# rownames(summary_df) <- row_names
+
+# # Rename the column names
+# col_names <- c("Term", "Estimate", "Std. Error", "t value", "Pr(>|t|)")
+# colnames(summary_df) <- col_names
+# 
+# # Print the summary table with kableExtra
+# library(kableExtra)
+# kable(summary_df, align = c("l", "r", "r", "r", "r"), 
+#       caption = "Summary of Linear Regression Model",) %>%
+#   kable_styling(bootstrap_options = "striped", full_width = FALSE)
+
+```
+```{r}
+#| label: fig-billssssss0
+#| fig-cap: More bills of penguins
+#| echo: false
+#| fig-width: 8
+#| fig-height: 4
+
+#level_order <- c(NA, "less than high school", "high school", "associate/junior college", "bachelor's", "graduate")
+
+tb1 <- cleaned_data %>% 
+  count(degree) %>% 
+  mutate(percent=percent(x=n/sum(n))) %>% 
+  rename(Degree = degree, Frequency = n, Proportion = percent)
+
+#tb1$Degree <- factor(tb1$Degree, levels = level_order)
+tb1<-tb1%>%arrange(factor(tb1$Degree,c("graduate", "bachelor's", "associate/junior college", "high school", "less than high school", NA)))
+
+  # arrange(Degree) %>% 
+  # mutate(Degree = fct_relevel(Degree, "less than high school", "high school", "associate/junior college", "bachelor's", "graduate", .missing = "NA"))
+
+tb1%>%kable(booktabs = TRUE)
+```
+
+```{r}
+#| label: fig-billssssss2
+#| fig-cap: More bills of penguins
+#| echo: false
+#| fig-width: 8
+#| fig-height: 4
+ggplot(data = cleaned_data, aes(x = educ)) +
+
+  geom_histogram(aes(y = ..density..), binwidth = 2, fill = "white") +
+  geom_density(alpha = .2, fill = "lightblue") +
+
+  # geom_histogram(bins = 15, fill = "lightblue", color = "white") +
+  labs(title = "Histogram of Respondents' Education", x = "Education (in Years)", y = "Desensity") +
+   scale_x_continuous(breaks = seq(0, 20, by = 1))+
+  theme_bw()
+```
+
+```{r}
+#| label: fig-billssssss3
+#| fig-cap: More bills of penguins
+#| echo: false
+#| fig-width: 8
+#| fig-height: 4
+ggplot(cleaned_data, aes(x = mapres10)) +
+  geom_histogram(aes(y = ..density.., fill = "mapres10"), alpha = 1, binwidth = 2,color = "black") +
+  geom_histogram(aes(x = papres10, y = ..density.., fill = "paeduc"), alpha = 0.4, binwidth = 2,color = "black") +
+  scale_fill_manual(values = c("mapres10" = "lightblue", "paeduc" = "red"), name = "Variables") +
+  labs(title = "Difference between Mother's and Father's Prestige Score",
+       x = "Prestige Score",
+       y = "Density")+
+  theme_bw()
+```
+
+```{r}
+#| label: fig-billssssss4
+#| fig-cap: More bills of penguins
+#| echo: false
+#| fig-width: 8
+#| fig-height: 4
+ggplot(cleaned_data, aes(x = maeduc)) +
+  geom_histogram(aes(y = ..density.., fill = "maeduc"), alpha = 1, binwidth = 1, color = "black") +
+  geom_histogram(aes(x = paeduc, y = ..density.., fill = "paeduc"), alpha = 0.4, binwidth = 1, color = "black") +
+  scale_fill_manual(values = c("maeduc" = "lightblue", "paeduc" = "red"), name = "Variables") +
+  labs(title = "Differene between Mother's Education and Father's Education",
+       x = "Education (in Years)",
+       y = "Density") +
+  theme_bw()
 ```
@@
 ```{r}
-#| label: fig-billssssss
+#| label: fig-billssssss5
 #| fig-cap: More bills of penguins
@@
 
-ggplot(raw_data, aes(x = papres10, y = educ)) +
-  geom_point(alpha=0.3)
+ggplot(cleaned_data, aes(x = educ, fill = factor(born))) +
+  geom_histogram(position = "dodge", binwidth = 1.5, alpha = 0.5, color = "black")+
+   labs(title = "Education and Birthplace (if outside of US) for the Respondent",
+       x = "Education (in Years)",
+       y = "Count",
+       fill = "Birthplace of the Respondent \n (if outside of US)")+
+  theme_bw()
+```
+
+```{r}
+#| label: fig-billssssss6
+#| fig-cap: More bills of penguins
+#| echo: false
+#| fig-width: 8
+#| fig-height: 4
+ggplot(cleaned_data, aes(x = educ,fill = factor(reg16))) +
+  geom_histogram(binwidth = 2, color = "black") +
+  facet_wrap(~ reg16, ncol = 3)+
+  labs(title = "Respondent's Living Region at Age of 16 v.s. Education",
+       x = "Education (in Years)",
+       y = "Count")+
+  theme_bw()+
+  theme(legend.position = "none")
+```
+
+```{r}
+#| label: fig-billssssss7
+#| fig-cap: More bills of penguins
+#| echo: false
+#| fig-width: 8
+#| fig-height: 4
+ggplot(cleaned_data, aes(x = maeduc, y = paeduc, color = factor(incom16))) +
+  geom_point() +
+  stat_smooth(method = "loess", se = FALSE) +
+  labs(x = "Mother's Education (in Years)", y = "Father's Education (in Years)", color = "Family Income")+
+  theme_bw()
 ```

### 2023-03-17 scripts/02-data_cleaning.R

@@ -21,7 +21,7 @@ raw_data <-
 
 raw_data <- 
   raw_data %>% 
-  select(mapres10,papres10,paeduc,maeduc,educ)
+  dplyr::select(mapres10,papres10,paeduc,maeduc,educ,mawrkslf,pawrkslf,degree,born,sex,reg16,incom16)
 
 #change class
 raw_data$mapres10 <- as.character(raw_data$mapres10)
@@ -30,30 +30,39 @@ raw_data$maeduc <- as.character(raw_data$maeduc)
 raw_data$paeduc <- as.character(raw_data$paeduc)
 raw_data$educ <- as.character(raw_data$educ)
 
-#change no formal to 0
-cleaned_data<-raw_data%>%mutate_all(~ifelse(. == "no formal schooling", 0, .))%>%
-  mutate_all(as.numeric)
+# Replace "no formal schooling" with 0 in the specified column
+raw_data$educ[raw_data$educ== "no formal schooling"] <- 0
+raw_data$maeduc[raw_data$maeduc== "no formal schooling"] <- 0
+raw_data$paeduc[raw_data$paeduc== "no formal schooling"] <- 0
+
+# Change to numeric
+raw_data$mapres10 <- as.numeric(raw_data$mapres10)
+raw_data$papres10 <- as.numeric(raw_data$papres10)
+raw_data$maeduc <- as.numeric(raw_data$maeduc)
+raw_data$paeduc <- as.numeric(raw_data$paeduc)
+raw_data$educ <- as.numeric(raw_data$educ)
 
 #group variables into levels
 
-mepa<-median(cleaned_data$papres10,na.rm=TRUE)
-mema<-median(cleaned_data$mapres10,na.rm=TRUE)
+mepa<-median(raw_data$papres10,na.rm=TRUE)
+mema<-median(raw_data$mapres10,na.rm=TRUE)
 
-cleaned_data <- cleaned_data %>%
+raw_data <- raw_data %>%
   mutate(father_factor = case_when(
     papres10 < mepa ~ "Lower than Median",
     papres10 >= mepa ~ "Above and Equal to Median"
   ))
 
-cleaned_data <- cleaned_data %>%
+raw_data <- raw_data %>%
   mutate(mother_factor = case_when(
     mapres10 < mema ~ "Lower than Median",
     mapres10 >= mema ~ "Above and Equal to Median"
   ))
 
+cleaned_data<-raw_data
 
 #### Save data ####
[…]

### 2023-03-18 outputs/paper/paper.qmd
add the text

@@
 You can and should cross-reference sections and sub-sections. For instance, @sec-data and @sec-first-point.
+
+In the past, family status variables, such as the socio-economic status and educational level of parents, were commonly considered as strong predictors of academic achievement in children.
+More specifically, parents who have lower levels of education, those with higher levels of education are more inclined to view higher education as desirable and encourage their children to excel academically.
+They also tend to hold higher expectations for their children's academic performance.
+(https://www.academia.edu/2712079/How_does_parents_education_level_influence_parenting_and_childrens_achievement)
+
+One potential explanation for this relationship is that higher levels of education can provide parents with access to resources such as income, time, energy, and community contacts, which can enable greater involvement in a child's education.
+As a result, the influence of family status variables on children's academic achievement may be best understood as a complex interaction between status and process variables (https://files.eric.ed.gov/fulltext/EJ1079955.pdf).
+
+However, recent research has indicated that the relationship between these factors and academic achievement is not always direct.
+Instead, socio-economic status and parents' education are part of a larger set of psychological and sociological variables that can impact children's educational outcomes (https://files.eric.ed.gov/fulltext/EJ1079955.pdf).
+
+This paper will test and explore how parents’ social-economics status can affect childrens’ education.
+The data that will be used in this paper comes from the US General Social Survey from National Opinion Research Center at the University of Chicago.
+The factor of socio-economics status is measured using the occupation prestige score.
+It is a measure used in social science research to assess the level of social status or prestige associated with a particular occupation or profession, which were developed by the National Opinion Research Center at the University of Chicago.
+As part of the GSS, respondents are asked to rate the prestige or social standing of various occupations on a scale of 1 to 90, with higher scores indicating greater prestige.
+Another variable that has been used is the education level, which is a number of variables to indicate the number of years the respondents have spent in school and college.
+(https://gss.norc.org/Documents/reports/methodological-reports/MR122%20Occupational%20Prestige.pdf)
+
+*MISSING*briefly discuss the results and application.
 
@@
 
-#level_order <- c(NA, "less than high school", "high school", "associate/junior college", "bachelor's", "graduate")
-
 tb1 <- cleaned_data %>% 
@@
 
-#tb1$Degree <- factor(tb1$Degree, levels = level_order)
 tb1<-tb1%>%arrange(factor(tb1$Degree,c("graduate", "bachelor's", "associate/junior college", "high school", "less than high school", NA)))
-
-  # arrange(Degree) %>% 
-  # mutate(Degree = fct_relevel(Degree, "less than high school", "high school", "associate/junior college", "bachelor's", "graduate", .missing = "NA"))
 
@@
 #| fig-height: 4
-ggplot(cleaned_data, aes(x = maeduc, y = paeduc, color = factor(incom16))) +
-  geom_point() +
-  stat_smooth(method = "loess", se = FALSE) +
-  labs(x = "Mother's Education (in Years)", y = "Father's Education (in Years)", color = "Family Income")+
-  theme_bw()
+
 ```
@@
 ## Weaknesses and next steps
+*problems with prestige score:*
+Subjectivity: Even measures of prestige that attempt to capture objective indicators of social status, such as income or educational attainment, are still subject to subjective interpretation.
+People's perceptions of what constitutes a high-prestige occupation may vary based on their own cultural, social, or personal biases.
 
-Weaknesses and next steps should also be included.
+Historical and cultural context: Measures of prestige are not fixed over time and may vary across different societies or historical periods.
+For example, certain occupations that were highly prestigious in the past, such as blacksmith or scribe, may be less highly regarded today.
+
+Heterogeneity within occupations: Different individuals within the same occupation may have different levels of prestige, depending on factors such as their level of experience, education, or specialization.
+
+Overemphasis on certain aspects of occupation: Some measures of prestige may place too much emphasis on certain aspects of occupation, such as income or educational requirements, while overlooking other important factors such as job security, working conditions, or level of autonomy.
+
+Lack of nuance: Prestige scores may provide a broad measure of social status and occupational major groups, but they may not capture the complexity and diversity of occupational experiences, especially for those in lower-prestige occupations.
+(https://www.researchgate.net/publication/236747304_A_Scale_of_Occupational_Prestige_in_Canada_Based_on_NOC_Major_Groups)
+
+*problems with missing data:*
+Reduced statistical power is one such issue, which lowers the likelihood of rejecting the null hypothesis when it is false.
+Missing data can also cause bias in parameter estimation.
+The representativeness of the sample can also be reduced due to missing data.
+The analysis of the study can become more complex when missing data is present.
+These problems can threaten the validity of the trials and lead to invalid conclusions.
+(https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3668100/)
 

### 2023-03-20 outputs/paper/paper.qmd
modified the setting for plots and tables

@@
 #| fig-height: 4
-
+ggplot(cleaned_data, aes(x = educ, y = maeduc, color = incom16)) +
+  geom_jitter() +
+  stat_smooth(method = "lm", se = FALSE) +
+  facet_wrap(~incom16)+
+  labs(title = "Education vs Mother's Education with respect to Family Income",
+       x = "Education (in Years)",
+       y = "Mother's Educaion (in Years)",
+       fill = "Family Income \n (at Age of 16)")+
+  theme_bw()
 ```

### 2023-03-20 outputs/paper/paper.qmd
add tests and simulate the data

--- 2023-03-18 outputs/paper/paper.qmd
@@
 ```{r,echo=FALSE}
+#| include: false
+#| warning: false
+#| message: false
 cleaned_data <- read_csv(
@@
 ```{r}
-#| label: fig-bills1
-#| fig-cap: Bills of penguins
+#| message: false
 #| echo: false
+#| warning: false
+#| label: tbl-2021-gender
+#| tbl-cap: Linear Model and its Summary Statistics
 
@@
 
-# # Rename the column names
-# col_names <- c("Term", "Estimate", "Std. Error", "t value", "Pr(>|t|)")
-# colnames(summary_df) <- col_names
-# 
-# # Print the summary table with kableExtra
-# library(kableExtra)
-# kable(summary_df, align = c("l", "r", "r", "r", "r"), 
-#       caption = "Summary of Linear Regression Model",) %>%
-#   kable_styling(bootstrap_options = "striped", full_width = FALSE)
+# Rename the column names
+col_names <- c("Term", "Estimate", "Std. Error", "t value", "Pr(>|t|)")
+colnames(summary_df) <- col_names
 
+# Print the summary table with kableExtra
+library(kableExtra)
+kable(summary_df, align = c("l", "r", "r", "r", "r"),
+      caption = "Summary of Linear Regression Model")
 ```
 ```{r}
-#| label: fig-billssssss0
-#| fig-cap: More bills of penguins
+#| message: false
 #| echo: false
-#| fig-width: 8
-#| fig-height: 4
+#| warning: false
+#| label: tbl-degree
+#| tbl-cap: Respondents' Degree in 2021 Survey
 
@@
 
-tb1%>%kable(booktabs = TRUE)
+tb1%>%kable()
 ```
@@
 ```{r}
-#| label: fig-billssssss2
-#| fig-cap: More bills of penguins
+#| label: fig-educ
+#| fig-cap: Histogram of Respondents' Education
 #| echo: false
+#| warning: false
+#| message: false
 #| fig-width: 8
@@
 ```{r}
-#| label: fig-billssssss3
-#| fig-cap: More bills of penguins
+#| message: false
 #| echo: false
+#| warning: false
+#| label: tbl-educationyears
+#| tbl-cap: Number of respondents by gender for 2021 survey
+tb0 <- cleaned_data %>% summarise(Min = min(cleaned_data$educ,na.rm=T),
+                              Median = median(cleaned_data$educ,na.rm=T),
+                              Max = max(cleaned_data$educ,na.rm = T),
+                              Mean = mean(cleaned_data$educ,na.rm=T))
+
+tb0%>%kable()
+```
+
+```{r}
+#| label: fig-parentpre
+#| fig-cap: Mother's and Father's Prestige Score
+#| echo: false
+#| warning: false
+#| message: false
 #| fig-width: 8
 #| fig-height: 4
+# ggplot(cleaned_data, aes(x = mapres10)) +
+#   geom_histogram(aes(y = ..density.., fill = "mapres10"), alpha = 1, binwidth = 2,color = "black") +
+#   geom_histogram(aes(x = papres10, y = ..density.., fill = "paeduc"), alpha = 0.8, binwidth = 2,color = "black") +
+#   scale_fill_manual(values = c("mapres10" = "lightblue", "paeduc" = "darkred"), name = "Variables") +
+#   labs(title = "Mother's and Father's Prestige Score",
+#        x = "Prestige Score",
+#        y = "Density")+
+#   theme_bw()
+
+##edit this
 ggplot(cleaned_data, aes(x = mapres10)) +
-  geom_histogram(aes(y = ..density.., fill = "mapres10"), alpha = 1, binwidth = 2,color = "black") +
-  geom_histogram(aes(x = papres10, y = ..density.., fill = "paeduc"), alpha = 0.4, binwidth = 2,color = "black") +
-  scale_fill_manual(values = c("mapres10" = "lightblue", "paeduc" = "red"), name = "Variables") +
-  labs(title = "Difference between Mother's and Father's Prestige Score",
+  geom_histogram(aes(y = ..density.., fill = "Mother's Prestige Score"), alpha = 1, binwidth = 1, color = "black")+
+  geom_histogram(aes(x = papres10, y = ..density.., fill = "Father's Prestige Score"), alpha = 0.8, binwidth = 1, color = "black") +
+  scale_fill_manual(values = c("Mother's Prestige Score" = "darkgreen", "Father's Prestige Score" = "lightgreen"), name = "Variables") +
+  labs(title = "Mother's and Father's Prestige Score",
        x = "Prestige Score",
-       y = "Density")+
+       y = "Density") +
   theme_bw()
@@
 ```{r}
-#| label: fig-billssssss4
-#| fig-cap: More bills of penguins
+#| label: fig-parentseduc
+#| fig-cap: Mother's Education and Father's Education Level
 #| echo: false
+#| warning: false
+#| message: false
 #| fig-width: 8
@@
 ggplot(cleaned_data, aes(x = maeduc)) +
-  geom_histogram(aes(y = ..density.., fill = "maeduc"), alpha = 1, binwidth = 1, color = "black") +
-  geom_histogram(aes(x = paeduc, y = ..density.., fill = "paeduc"), alpha = 0.4, binwidth = 1, color = "black") +
-  scale_fill_manual(values = c("maeduc" = "lightblue", "paeduc" = "red"), name = "Variables") +
-  labs(title = "Differene between Mother's Education and Father's Education",
+geom_histogram(aes(y = ..density.., fill = "Mother's Education"), alpha = 1, binwidth = 0.5, color = "black") +
+geom_histogram(aes(x = paeduc, y = ..density.., fill = "Father's Education"), alpha = 0.8, binwidth = 0.5, color = "black") +
+scale_fill_manual(values = c("Mother's Education" = "darkgreen", "Father's Education" = "lightgreen"), name = "Variables") +
+labs(title = "Mother's Education and Father's Education Level",
 x = "Education (in Years)",
@@
 ```{r}
-#| label: fig-billssssss5
-#| fig-cap: More bills of penguins
+#| label: fig-birth
+#| fig-cap: Birthplace of the Respondent \n (if outside of US)
 #| echo: false
+#| warning: false
+#| message: false
 #| fig-width: 8
@@
 ```{r}
-#| label: fig-billssssss6
-#| fig-cap: More bills of penguins
+#| label: fig-region
+#| fig-cap: Respondent's Living Region at Age of 16 v.s. Education
 #| echo: false
+#| warning: false
+#| message: false
 #| fig-width: 8
@@
 ```{r}
-#| label: fig-billssssss7
-#| fig-cap: More bills of penguins
+#| label: fig-maincome
+#| fig-cap: Education vs Mother's Education with respect to Family Income
 #| echo: false
+#| warning: false
+#| message: false
 #| fig-width: 8
 #| fig-height: 4
+ggplot(cleaned_data, aes(x = educ, y = maeduc, color = incom16)) +
+  geom_jitter() +
+  stat_smooth(method = "lm", se = FALSE) +
+  facet_wrap(~incom16)+
+  labs(title = "Education vs Mother's Education with respect to Family Income",
+       x = "Education (in Years)",
+       y = "Mother's Educaion (in Years)",
+       color = "Family Income \n (at Age of 16)") +
+  scale_color_discrete(name = "Family Income \n (at Age of 16)") +
+  theme_bw()
+```
 
+```{r}
+#| label: fig-paincome
+#| fig-cap: Education vs Father's Education with respect to Family Income
+#| echo: false
+#| warning: false
+#| message: false
+#| fig-width: 8
+#| fig-height: 4
+ggplot(cleaned_data, aes(x = educ, y = paeduc, color = incom16)) +
+  geom_jitter() +
+  stat_smooth(method = "lm", se = FALSE) +
+  facet_wrap(~incom16)+
+  labs(title = "Education vs Father's Education with respect to Family Income",
+       x = "Education (in Years)",
+       y = "Father's Educaion (in Years)",
+  color = "Family Income \n (at Age of 16)") +
+  scale_color_discrete(name = "Family Income \n (at Age of 16)") +
+  theme_bw()
 ```
@@
 
-## Weaknesses and next steps
+## Weaknesses and next steps {#sec-weaknesses}
 *problems with prestige score:*

### 2023-03-20 scripts/00-simulate_data.R

@@ -1,19 +1,53 @@
 #### Preamble ####
-# Purpose: Simulates... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Data: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Data cleaning for GSS 2021
+# Author: Ruibo Sun
+# Data: 16 March 2023
+# Contact: ruibo.sun@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: Download all the packges under the Workspace section
+# Any other information needed? None
 
 
 #### Workspace setup ####
 library(tidyverse)
-# [...UPDATE THIS...]
+
 
 #### Simulate data ####
-# [...ADD CODE HERE...]
+set.seed(123) # set seed for reproducibility
+
+# generate 100 observations
+n <- 100 
+
+# generate parents' education level, with higher levels of education having a greater probability of occurrence
+parent_edu <- sample(c("less than high school", "high school", "some college", "bachelor's degree", "graduate degree"), 
+                     n, 
+                     replace = TRUE)
+
+# generate a binary variable indicating whether parents have higher education or not
+parent_higher_edu <- ifelse(parent_edu %in% c("bachelor's degree", "graduate degree"), 1, 0)
+
+# generate children's education, with higher probability of occurrence if parents have higher education
+child_edu <- sample(c("less than high school", "high school", "some college", "bachelor's degree", "graduate degree"), 
+                    n, 
+                    replace = TRUE)
+
+# generate a binary variable indicating whether children have higher education or not
+child_higher_edu <- ifelse(child_edu %in% c("bachelor's degree", "graduate degree"), 1, 0)
+
+# generate a variable indicating region of living during teenage age, with equal probability of occurrence
+region <- sample(c("North", "South", "East", "West"), n, replace = TRUE)
+
+# generate a random prestige score for parents, with higher education level and living in certain regions associated with higher scores
+prestige_score <- ifelse(parent_edu %in% c("bachelor's degree", "graduate degree"), 
+                         rnorm(n, mean = 75, sd = 15), 
+                         rnorm(n, mean = 50, sd = 10))
[…]

### 2023-03-20 scripts/04-test_data.R

@@ -1,15 +1,47 @@
 #### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Data: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Data cleaning for GSS 2021
+# Author: Ruibo Sun
+# Data: 16 March 2023
+# Contact: ruibo.sun@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: Download all the packges under the Workspace section
+# Any other information needed? None
 
 
 #### Workspace setup ####
 library(tidyverse)
-# [...UPDATE THIS...]
+cleaned_data <- read_csv(
+  file = here("inputs/data/GSS2021 2.csv"))
 
 #### Test data ####
+cleaned_data$mapres10|> class() == "numeric"
+
+cleaned_data$papres10|> class() == "numeric"
+
+cleaned_data$paeduc|> class() == "numeric"
+
+cleaned_data$maeduc|> class() == "numeric"
+
+cleaned_data$educ|> class() == "numeric"
+
+cleaned_data$mawrkslf|> class() == "character"
+
+cleaned_data$pawrkslf|> class() == "character"
+
+cleaned_data$degree|> class() == "character"
+
+cleaned_data$born|> class() == "character"
+
+cleaned_data$sex|> class() == "character"
+
+cleaned_data$degree|> class() == "character"
+
+cleaned_data$incom16|> class() == "character"
+
+cleaned_data$reg16|> class() == "character"
+
+cleaned_data$maeduc|> min(na.rm = T) >= 0
+
+cleaned_data$paeduc|>min(na.rm = T) >= 0
+
+cleaned_data$educ|>min(na.rm = T) >= 0

### 2023-03-21 outputs/paper/paper.qmd
small changes in the codes for plot

@@
 number-sections: true
-#bibliography: references.bib
+bibliography: references.bib
 ---
@@
 library(forcats)
+library(patchwork)
 ```
@@
 
+# Results
+
 ```{r}
@@
 
-# Get the summary statistics and tidy them up
-tidy_summary <- tidy(summary(model))
+modelsummary::modelsummary(
+  list(
+    "Children's Education" = model
+  ),
+  fmt = 2,
+  statistic = "conf.int"
+)
+# 
+# # Get the summary statistics and tidy them up
+# tidy_summary <- tidy(summary(model))
+# 
+# # Create a data frame for the tidy summary
+# summary_df <- as.data.frame(tidy_summary)
+# 
+# # # Rename the row names
+# # row_names <- c("Intercept", "Mother's Prestiges Score", "Father's Prestiges Score", "Father's Education", "Mother's Education")
+# # rownames(summary_df) <- row_names
+# 
+# # Rename the column names
+# col_names <- c("Term", "Estimate", "Std. Error", "t value", "Pr(>|t|)")
+# colnames(summary_df) <- col_names
+# 
+# # Print the summary table with kableExtra
+# 
+# kable(summary_df, align = c("l", "r", "r", "r", "r"),
+#       caption = "Summary of Linear Regression Model")
+```
 
-# Create a data frame for the tidy summary
-summary_df <- as.data.frame(tidy_summary)
-
-# # Rename the row names
-# row_names <- c("Intercept", "Mother's Prestiges Score", "Father's Prestiges Score", "Father's Education", "Mother's Education")
-# rownames(summary_df) <- row_names
-
-# Rename the column names
-col_names <- c("Term", "Estimate", "Std. Error", "t value", "Pr(>|t|)")
-colnames(summary_df) <- col_names
-
-# Print the summary table with kableExtra
-library(kableExtra)
-kable(summary_df, align = c("l", "r", "r", "r", "r"),
-      caption = "Summary of Linear Regression Model")
-```
 ```{r}
@@
   labs(title = "Histogram of Respondents' Education", x = "Education (in Years)", y = "Desensity") +
-   scale_x_continuous(breaks = seq(0, 20, by = 1))+
+  scale_x_continuous(breaks = seq(0, 20, by = 1)) +
   theme_bw()
@@
 
-tb0%>%kable()
+tb0 %>% kable()
 ```
@@
 
-##edit this
+#prestige score distribution
 ggplot(cleaned_data, aes(x = mapres10)) +
@@
   theme_bw()
+```
+
+```{r}
+#| label: fig-presandeduc
+#| fig-cap: Mother's and Father's Prestige Score
+#| echo: false
+#| warning: false
+#| message: false
+#| fig-width: 8
+#| fig-height: 4
+#pres vs. educ
+p1 <- ggplot(cleaned_data, aes(x=educ, y=papres10)) +
+  geom_point(color="#69b3a2", size=2) +
+  ggtitle("Father's Prestige Score vs Children's Education") +
+  labs(x="Education (in Years)", y="Father's Prestige Score")+
+  theme_bw()
+p2 <- ggplot(cleaned_data, aes(x=educ, y=mapres10)) +
+  geom_point(color="darkgreen",size=2) +
+  ggtitle("Mother's Prestige Score vs Children's Education") +
+  labs(x="Education (in Years)", y="Mother's Prestige Score")+
+  theme_bw()
+
+p1 + p2
 ```
@@
 #| label: fig-birth
-#| fig-cap: Birthplace of the Respondent \n (if outside of US)
+#| fig-cap: Birthplace of the Respondent (if outside of US)
 #| echo: false
@@
        y = "Count",
-       fill = "Birthplace of the Respondent \n (if outside of US)")+
+       fill = "Birthplace of the Respondent (if outside of US)")+
   theme_bw()
@@
   geom_histogram(binwidth = 2, color = "black") +
-  facet_wrap(~ reg16, ncol = 3)+
+  facet_wrap(~ reg16, ncol = 2)+
   labs(title = "Respondent's Living Region at Age of 16 v.s. Education",
@@
        y = "Mother's Educaion (in Years)",
-       color = "Family Income \n (at Age of 16)") +
-  scale_color_discrete(name = "Family Income \n (at Age of 16)") +
+       color = "Family Income (at Age of 16)") +
+  scale_color_discrete(name = "Family Income (at Age of 16)") +
   theme_bw()
@@
 #| fig-height: 4
+
 ggplot(cleaned_data, aes(x = educ, y = paeduc, color = incom16)) +
@@
        y = "Father's Educaion (in Years)",
-  color = "Family Income \n (at Age of 16)") +
-  scale_color_discrete(name = "Family Income \n (at Age of 16)") +
+  color = "Family Income (at Age of 16)") +
+  scale_color_discrete(name = "Family Income (at Age of 16)") +
   theme_bw()
@@
 
-# Results
-
 # Discussion
@@
 *problems with prestige score:*
+
 Subjectivity: Even measures of prestige that attempt to capture objective indicators of social status, such as income or educational attainment, are still subject to subjective interpretation.
@@
 *problems with missing data:*
+
 Reduced statistical power is one such issue, which lowers the likelihood of rejecting the null hypothesis when it is false.

### 2023-03-24 outputs/paper/paper.qmd
changes in files and folder

@@
 p1 <- ggplot(cleaned_data, aes(x=educ, y=papres10)) +
-  geom_point(color="#69b3a2", size=2) +
+  geom_point(color="#69b3a2", size=2,alpha=0.3) +
   ggtitle("Father's Prestige Score vs Children's Education") +
@@
 p2 <- ggplot(cleaned_data, aes(x=educ, y=mapres10)) +
-  geom_point(color="darkgreen",size=2) +
+  geom_point(color="darkgreen",size=2,alpha=0.3) +
   ggtitle("Mother's Prestige Score vs Children's Education") +
@@
        y = "Count",
-       fill = "Birthplace of the Respondent (if outside of US)")+
+       fill = "Birthplace of the Respondent (if in, of US)")+
   theme_bw()
+```
+```{r,echo=FALSE}
+tb3 <- cleaned_data %>% 
+  count(born) %>% 
+  mutate(percent=percent(x=n/sum(n)))
 ```
@@
 
+```{r,echo=FALSE}
+tb4 <- cleaned_data %>% 
+  count(reg16) %>% 
+  mutate(percent=percent(x=n/sum(n)))
+```
+
 ```{r}
@@
 ggplot(cleaned_data, aes(x = educ, y = maeduc, color = incom16)) +
-  geom_jitter() +
+  geom_jitter(alpha = 0.2) +
   stat_smooth(method = "lm", se = FALSE) +
@@
 ggplot(cleaned_data, aes(x = educ, y = paeduc, color = incom16)) +
-  geom_jitter() +
+  geom_jitter(alpha = 0.2) +
   stat_smooth(method = "lm", se = FALSE) +
@@
 
-# Model
+# Model {#sec-model}
 

### 2023-03-24 outputs/paper/paper.qmd
remove all the title from the plots

--- 2023-03-21 outputs/paper/paper.qmd
@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
+title: "The Relationship between Family Status and Academic Achievement in Children"
 author: 
@@
 date-format: "12-03-2023"
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "This paper explores the relationship between family status variables, such as parents' socio-economic status and educational level, and their impact on children's academic achievement and education level using the data from 2021 GSS.
+Previous studies have suggested that parents' education can positively influence their children's academic performance due to increased access to resources and greater involvement in their education.
+However, recent research has demonstrated that this relationship is complex and can be influenced by a variety of psychological and sociological factors.
+The findings suggest that parents' education, family income, and prestige score all have a positive correlation with children's education.
+Further research is needed to fully understand the complexities of this relationship and use the results to improve educational policies and programs."
 format: pdf
@@
 library(scales)
-library(dplyr)
-library(purrr)
-library(forcats)
+#library(purrr)
+#library(forcats)
 library(patchwork)
@@
 #| warning: false
-#| label: tbl-2021-gender
-#| tbl-cap: Linear Model and its Summary Statistics
-
-# Fit the model
-model <- lm(educ ~ mapres10 + papres10 + paeduc + maeduc, data = cleaned_data)
-
-modelsummary::modelsummary(
-  list(
-    "Children's Education" = model
-  ),
-  fmt = 2,
-  statistic = "conf.int"
-)
-# 
-# # Get the summary statistics and tidy them up
-# tidy_summary <- tidy(summary(model))
-# 
-# # Create a data frame for the tidy summary
-# summary_df <- as.data.frame(tidy_summary)
-# 
-# # # Rename the row names
-# # row_names <- c("Intercept", "Mother's Prestiges Score", "Father's Prestiges Score", "Father's Education", "Mother's Education")
-# # rownames(summary_df) <- row_names
-# 
-# # Rename the column names
-# col_names <- c("Term", "Estimate", "Std. Error", "t value", "Pr(>|t|)")
-# colnames(summary_df) <- col_names
-# 
-# # Print the summary table with kableExtra
-# 
-# kable(summary_df, align = c("l", "r", "r", "r", "r"),
-#       caption = "Summary of Linear Regression Model")
-```
-
-```{r}
-#| message: false
-#| echo: false
-#| warning: false
 #| label: tbl-degree
@@
   # geom_histogram(bins = 15, fill = "lightblue", color = "white") +
-  labs(title = "Histogram of Respondents' Education", x = "Education (in Years)", y = "Desensity") +
+  labs(x = "Education (in Years)", y = "Desensity") +
   scale_x_continuous(breaks = seq(0, 20, by = 1)) +
@@
   scale_fill_manual(values = c("Mother's Prestige Score" = "darkgreen", "Father's Prestige Score" = "lightgreen"), name = "Variables") +
-  labs(title = "Mother's and Father's Prestige Score",
-       x = "Prestige Score",
+  labs(x = "Prestige Score",
        y = "Density") +
@@
 p1 <- ggplot(cleaned_data, aes(x=educ, y=papres10)) +
-  geom_point(color="#69b3a2", size=2) +
-  ggtitle("Father's Prestige Score vs Children's Education") +
+  geom_point(color="#69b3a2", size=2,alpha=0.3) +
   labs(x="Education (in Years)", y="Father's Prestige Score")+
@@
 p2 <- ggplot(cleaned_data, aes(x=educ, y=mapres10)) +
-  geom_point(color="darkgreen",size=2) +
-  ggtitle("Mother's Prestige Score vs Children's Education") +
+  geom_point(color="darkgreen",size=2,alpha=0.3) +
   labs(x="Education (in Years)", y="Mother's Prestige Score")+
@@
 scale_fill_manual(values = c("Mother's Education" = "darkgreen", "Father's Education" = "lightgreen"), name = "Variables") +
-labs(title = "Mother's Education and Father's Education Level",
-x = "Education (in Years)",
+labs(x = "Education (in Years)",
 y = "Density") +
@@
   geom_histogram(position = "dodge", binwidth = 1.5, alpha = 0.5, color = "black")+
-   labs(title = "Education and Birthplace (if outside of US) for the Respondent",
-       x = "Education (in Years)",
+   labs(x = "Education (in Years)",
        y = "Count",
-       fill = "Birthplace of the Respondent (if outside of US)")+
+       fill = "Birthplace of the Respondent (if in, of US)")+
   theme_bw()
+```
+```{r,echo=FALSE}
+tb3 <- cleaned_data %>% 
+  count(born) %>% 
+  mutate(percent=percent(x=n/sum(n)))
 ```
@@
   facet_wrap(~ reg16, ncol = 2)+
-  labs(title = "Respondent's Living Region at Age of 16 v.s. Education",
-       x = "Education (in Years)",
+  labs(x = "Education (in Years)",
        y = "Count")+
@@
   theme(legend.position = "none")
+```
+
+```{r,echo=FALSE}
+tb4 <- cleaned_data %>% 
+  count(reg16) %>% 
+  mutate(percent=percent(x=n/sum(n)))
 ```
@@
 ggplot(cleaned_data, aes(x = educ, y = maeduc, color = incom16)) +
-  geom_jitter() +
+  geom_jitter(alpha = 0.2) +
   stat_smooth(method = "lm", se = FALSE) +
   facet_wrap(~incom16)+
-  labs(title = "Education vs Mother's Education with respect to Family Income",
-       x = "Education (in Years)",
+  labs(x = "Education (in Years)",
        y = "Mother's Educaion (in Years)",
@@
 ggplot(cleaned_data, aes(x = educ, y = paeduc, color = incom16)) +
-  geom_jitter() +
+  geom_jitter(alpha = 0.2) +
   stat_smooth(method = "lm", se = FALSE) +
   facet_wrap(~incom16)+
-  labs(title = "Education vs Father's Education with respect to Family Income",
-       x = "Education (in Years)",
+  labs(x = "Education (in Years)",
        y = "Father's Educaion (in Years)",
@@
 
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+#| label: tbl-lmeduc
+#| tbl-cap: Linear Model and its Summary Statistics
+
+# Fit the model
+model <- lm(educ ~ mapres10 + papres10 + paeduc + maeduc, data = cleaned_data)
+
+modelsummary::modelsummary(
+  list(
+    "Children's Education" = model
+  ),
+  fmt = 2,
+  statistic = "conf.int"
+)
+```
+
 Talk way more about it.
 
-# Model
+# Model {#sec-model}
 

### 2023-03-24 outputs/paper/paper.qmd
second version of the draft

@@
 
-You can and should cross-reference sections and sub-sections. For instance, @sec-data and @sec-first-point.
-
 In the past, family status variables, such as the socio-economic status and educational level of parents, were commonly considered as strong predictors of academic achievement in children.
-More specifically, parents who have lower levels of education, those with higher levels of education are more inclined to view higher education as desirable and encourage their children to excel academically.
-They also tend to hold higher expectations for their children's academic performance.
-(https://www.academia.edu/2712079/How_does_parents_education_level_influence_parenting_and_childrens_achievement)
+More specifically, parents who have lower levels of education, and those with higher levels of education are more inclined to view higher education as desirable and encourage their children to excel academically.
+They also tend to hold higher expectations for their children's academic performance [@davis_2014].
 
 One potential explanation for this relationship is that higher levels of education can provide parents with access to resources such as income, time, energy, and community contacts, which can enable greater involvement in a child's education.
-As a result, the influence of family status variables on children's academic achievement may be best understood as a complex interaction between status and process variables (https://files.eric.ed.gov/fulltext/EJ1079955.pdf).
+As a result, the influence of family status variables on children's academic achievement may be best understood as a complex interaction between status and process variables [@khan].
 
 However, recent research has indicated that the relationship between these factors and academic achievement is not always direct.
-Instead, socio-economic status and parents' education are part of a larger set of psychological and sociological variables that can impact children's educational outcomes (https://files.eric.ed.gov/fulltext/EJ1079955.pdf).
+Instead, socio-economic status and parents' education are part of a larger set of psychological and sociological variables that can impact children's educational outcomes [@khan].
 
-This paper will test and explore how parents’ social-economics status can affect childrens’ education.
-The data that will be used in this paper comes from the US General Social Survey from National Opinion Research Center at the University of Chicago.
-The factor of socio-economics status is measured using the occupation prestige score.
-It is a measure used in social science research to assess the level of social status or prestige associated with a particular occupation or profession, which were developed by the National Opinion Research Center at the University of Chicago.
-As part of the GSS, respondents are asked to rate the prestige or social standing of various occupations on a scale of 1 to 90, with higher scores indicating greater prestige.
-Another variable that has been used is the education level, which is a number of variables to indicate the number of years the respondents have spent in school and college.
-(https://gss.norc.org/Documents/reports/methodological-reports/MR122%20Occupational%20Prestige.pdf)
+In short, the findings in this paper demonstrate that parents’ education has a positive correlation with children’s education.
+Additionally, the region of living during the teenage years can also impact children’s education.
+Parents’ prestige score seems to be another factor in children’s education.
+However, there is some limitation on the definition of prestige score that may influence the overall results.
+This will be discussed in the @sec-weakness.
 
-*MISSING*briefly discuss the results and application.
+The data will be presented clearly and succinctly with plots and tables.
+This following analysis is processed in R [@R]  with packages of tidyverse [@tidyverse], dplyr [@dplyr], here [@here], haven [@haven] and broom [@broom].
+The tables are constructed via knitr [@knitr], scales [@scales] and kableExtra [@kableExtra].
+The package inside tidyverse helps to create the plots in ggplot2 [@tidyverse].
+This paper is knitted as a PDF file by the packages of R markdown [@RMarkdown] and formatted using patchwork [@patchwork].
 
@@
 
-Our data is of penguins (@fig-bills).
+## Source
+The 2021 General Social Survey (GSS) is a nationally representative survey conducted to collect data on social trends and attitudes among people living in the United States.
+Data was collected through face-to-face interviews with adult residents, covering a wide range of variables that are of interest to social scientists, policymakers, and the general public.
+The variables measured in the survey include demographics, employment, education, health, family, and social attitudes [@gss_data_explorer].
+
+The target population is US adults aged 18 and older.
+The GSS uses a multistage probability sampling approach to recruit its sample from the US Census Bureau's Master Address File.
+One strength of the GSS is its long history and large sample size, but a potential weakness is a reliance on self-reported data, which can be subject to response bias.
+The questionnaire includes both closed-ended and open-ended questions, but there may be concerns about biased or leading questions.
+The GSS employs various methods to adjust for non-response and ensure the representativeness of the sample.
+One strength of the GSS is its long history and large sample size, but a potential weakness is a reliance on self-reported data, which can be subject to response bias.
+The questionnaire includes both closed-ended and open-ended questions, but there may be concerns about biased or leading questions.
+[@norcuofch] The GSS employs various methods to adjust for non-response and ensure the representativeness of the sample.
+The limitation will be explained in detail in the @sec-weakness.
+
+This paper will test and explore how parents’ social-economics status can affect children’s education.
+The data that will be used in this paper comes from the US General Social Survey from the National Opinion Research Center at the University of Chicago.
+The factor of socio-economics status is measured using the occupation prestige score.
+It is a measure used in social science research to assess the level of social status or prestige associated with a particular occupation or profession, which was developed by the National Opinion Research Center at the University of Chicago.
+As part of the GSS, respondents are asked to rate the prestige or social standing of various occupations on a scale of 1 to 90, with higher scores indicating greater prestige.
+[@smith] Another variable that has been used is the education level, which is a number of variables to indicate the number of years the respondents have spent in school and college.
+Other variables will be discussed in @sec-data.
+
+## Data cleaning
+
+For simplicity purposes, only variables that are closely related to the topics are selected for further analysis.
+These variables are mapres10, papres10, paeduc, maeduc, educ, mawrkslf, pawrkslf, degree, born, sex, reg16, and incom16.
+
+- **mapres10**: A numeric variable indicating the respondent's mother’s prestige score.
+- **papres10**: A numeric variable indicating the respondent's father’s prestige score.
+- **paeduc**: A numeric variable indicating the number of years of education completed by the respondent's father.
+- **maeduc**: A numeric variable indicating the number of years of education completed by the respondent's mother.
+- **educ**: A numeric variable indicating the number of years of education completed by the respondent.
+- **mawrkslf**: A categorical variable indicating whether the respondent's mother was employed for wages or salary.
+- **pawrkslf**: A categorical variable indicating whether the respondent's father was employed for wages or salary.
+- **degree**: A categorical variable indicating the highest degree earned by the respondent.
+- **born**: A categorical variable indicating the region of the US where the respondent was born.
+- **sex**: A categorical variable indicating the respondent's sex.
+- **reg16**: A categorical variable indicating the region of the US where the respondent lived at age of 16.
+- **incom16**: A categorical variable indicating the respondent's total household at age of 16.
+
+The level of “no formal schooling” in **maeduc**, **paeduc** and **educ** is modified as the integer zero to make the variables consistent.
+All the numeric variables are shown as integers, but their class are in terms of character in R.
+Hence, the variables of **paeduc**, **maeduc**, **educ**, **mapres10** and **papres10** have been converted into a class of numeric for further analysis.
+
+None of the missing values (NA) is removed in this paper.
+The reason is that removing the missing values in every column would result in a huge loss in the dataset.
+To maintain the representativeness of the data, missing values are automatically filtered by R, rather than removed manually.
+The possible limitation and effect of the missing values will be discussed in the @sec-weakness.
 
@@
 #| label: tbl-degree
-#| tbl-cap: Respondents' Degree in 2021 Survey
-
-tb1 <- cleaned_data %>% 
-  count(degree) %>% 
-  mutate(percent=percent(x=n/sum(n))) %>% 
-  rename(Degree = degree, Frequency = n, Proportion = percent)
-
-tb1<-tb1%>%arrange(factor(tb1$Degree,c("graduate", "bachelor's", "associate/junior college", "high school", "less than high school", NA)))
-
-tb1%>%kable()
+#| tbl-cap: Number of respondents by degree for 2021 survey
+tb1 <- cleaned_data %>% count(degree) %>%
+  mutate(percent=percent(x=n/sum(n)))
+tb1<-tb1%>%arrange(factor(tb1$degree,c("graduate", "bachelor's", "associate/junior college", "high school", "less than high school", NA)))
+colnames(tb1) <- c("Degree", "Total", "Proportion")
+tb1%>%kable(booktabs = TRUE)
 ```
+The respondents in this study are the children and their parents' education has also been collected in the GSS.
+As for the response variables, children’s education level, there are two ways to measure.
+One is the degree, and the other one is the number of years in school.
+In @tbl-degree, it shows the distribution of education in terms of degrees.
+The data reveals that the most common degree category among the group is high school, with 1597 individuals (39.6%) having this level of education.
+Following this, the next most common category is a bachelor's degree, with 1036 individuals (25.7%) having completed this level of education.
+The data also indicates that a graduate degree is the third most common category, with 760 individuals (18.8%) having this level of education.
+Associate/junior college degree is less common than a bachelor's or graduate degree, with 370 individuals (9.2%) having completed this level of education.
+However, a smaller proportion of individuals (6.1%) have less than a high school education.
+Additionally, there are also 23 individuals (0.6%) whose educational degree information is not available (NA).
+Overall, the data provide insights into the educational distribution of the given group and may be useful for drawing conclusions and making informed decisions related to education.
 
@@
 ```
+The second variable utilized to quantify education is relation to the number of years of schooling.
+In @fig-educ, this histogram shows an overall distribution of the respondents' education in years.
+It is clear that the distribution is right-skewed, which indicates that there are more respondents with lower levels of education than with higher levels of education.
+In other words, the bulk of the respondents are clustered towards the left side of the histogram (lower levels of education), and there are relatively fewer respondents towards the right side of the histogram (higher levels of education).
+In short, the majority of respondents in 2021 GSS are highly educated.
 
@@
 #| warning: false
-#| label: tbl-educationyears
-#| tbl-cap: Number of respondents by gender for 2021 survey
+#| label: tbl-educ
+#| tbl-cap: Number of respondents by education for 2021 survey
 tb0 <- cleaned_data %>% summarise(Min = min(cleaned_data$educ,na.rm=T),
@@
 ```
+To see a more accurate numerical summary, @tbl-educ is created.
+In the dataset labelled @sec-data, the category denoting "no formal schooling" has been redefined as an integer value of 0.
+According to the information presented in the @tbl-educ table, the range of years of schooling reported by respondents spans from 0 to 20 years.
+The median number of years of schooling reported is 15, and the mean is 14.76904.
+This mean value roughly corresponds to three years in college.
+Taken together, these findings suggest that the respondents surveyed in the 2021 GSS have generally completed their high school education and are highly educated.
 
@@
 ```
+In @fig-parentpre, it shows the distribution of respondent’s father’s and mother’s prestige scores.
+The distribution of father’s prestige score and the mother’s prestige score have quite different distributions.
 
@@
 #| label: fig-presandeduc
-#| fig-cap: Mother's and Father's Prestige Score
+#| fig-cap: Mother's and Father's Prestige Score with children's education
 #| echo: false
@@
 ```
+In order to examine the relationship between parents’ prestige scores and children’s education level.
+The next plot, @fig-presandeduc, is used to show the relation.
+These two scatter plots show the father’s prestige score and the mother’s prestige score with their children’s education.
+It is hard to see a clear relation or pattern based on these two scatter plots.
+To verify if the prestige score is one of the factors, a linear model will be used in the @sec-model.
 
@@
 ```
-
-Talk more about it.
-
-Also bills and their average (@fig-billssssss).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work.)
+Another key variable is the parent’s education level.
+In @fig-parentseduc, it shows the distribution of the mother’s education and the father’s education.
+One interesting result is that the distribution of education for mother and father is similar, or even almost identical.
+This indicates that there is a strong correlation between the education levels of mothers and fathers.
+In other words, if a mother has a high level of education, it is likely that the father also has a high level of education, and vice versa.
 
@@
 #| label: fig-birth
-#| fig-cap: Birthplace of the Respondent (if outside of US)
+#| fig-cap: Birthplace of the Respondent (if in US)
 #| echo: false
@@
        y = "Count",
-       fill = "Birthplace of the Respondent (if in, of US)")+
+       fill = "Birthplace of the Respondent (if in US)")+
   theme_bw()
 ```
+Based on [@overman], the birthplace may have an impact on children’s education.
+In @fig-birth, it is used to test the relationship between birthplace and children’s education.
+Note that the question of birthplace only asks if the respondent was born in the US or not.
+In general, most of the respondents have received at least 10 years of education despite their birthplace.
+However, the bar for respondents who were born outside of the US is much higher than the other groups.
+This is because 87.2% of the respondents were born in the US, and only 11.0% answered “no”.
+Additionally, there are 72 NA, which take a proportion of 1.8%.
+Given the huge differences in the sample sizes between groups, it is hard to draw a conclusion of whether birthplace may affect the education level.
+
 ```{r,echo=FALSE}
@@
 ```
+According to the study [@nieuwenhuis], there appears to be an association between neighborhood and education.
+In the 2021 GSS, the neighbourhood variable is being measured by the living region at the age of 16.
+The distribution for all regions, as shown in @fig-region, follows a similar bimodal shape, with the exception of the NA group.
+This could indicate that the region of living has a minimal or no impact on children's education levels.
+However, it is important to note that the measurement of the neighborhood through living regions at age 16 may not fully capture the impact of the neighborhood on education, which will be discussed in the @sec-weakness.
 
@@
 
+To examine the relationship between family income and education, the @fig-maincome and @fig-paincome visualize this relation.
+In @fig-maincome, the vertical axis represents the mother’s education and the horizontal axis represents the children’s education.
+The respondents were asked to rate their family income at age of 16 in the following category: (1) far above average, (2) above average, (3) average, (4) below average and (5) far below average.
+Due to missing values, the sixth group is NA, which is coloured grey.
+Overall, mothers’ education and children’s education are positively correlated despite the level of income.
+To be more specific, for respondents whose family income was far above average, the slope seems to be steeper than the other groups, which indicates that for those with higher family income, there may be a stronger relationship between the education of the mother and the education of the children.
+
 ```{r}
@@
 ```
+In @fig-paincome, the income groups are the same as before.
+However, the plot is now examining the correlation between the father’s education and the children’s education.
+Once again, there is a positive correlation between the education level of fathers and their children's education, regardless of family income.
+However, for respondents who reported a "far above average" family income, the relationship between the father's education and children's education appears to be stronger than for other income groups.
+This suggests that higher family income may be associated with a more pronounced link between paternal and children's education levels.
+Comparing the two plots, the slopes in @fig-paincome for all income groups are much steeper than the slopes in @fig-maincome.
+This may suggest that there is a stronger relationship between fathers' education and children's education compared to mothers' education and children's education given the same income level.
+
+# Model {#sec-model}
+
+$$
+\hat{Y}=\hat{\beta_0}+\hat{\beta_1}X_{mapres10}+\hat{\beta_2}X_{papres10}+\hat{\beta_3}X_{paeduc}+\hat{\beta_4}X_{maeduc}
+$$ {#eq-model}
 
@@
 ```
+The @tbl-lmeduc is built based on the @eq-model:
 
-Talk way more about it.
+- The intercept is 11.32, which means that when all the independent variables are zero, the dependent variable has a value of 11.32.
+- The coefficients for the independent variables mapres10 and papres10 are both 0.01, which means that a one-unit increase in either of these variables is associated with an increase of 0.01 in the dependent variable.
+- The coefficients for the independent variables paeduc and maeduc are 0.18 and 0.05, respectively.
+This means that a one-unit increase in paeduc is associated with an increase of 0.18 in the dependent variable, and a one-unit increase in maeduc is associated with an increase of 0.05 in the dependent variable.
+- The number of observations used in the model is 1980.
+The R-squared value is 0.124, which means that the independent variables explain 12.4% of the variation in the dependent variable.
+- The adjusted R-squared value is 0.122.
+- The AIC and BIC values are measures of the goodness of fit of the model. Lower values indicate a better fit.
+- The log-likelihood is a measure of how well the model fits the data.
+- The RMSE value is 2.44, which is the root mean squared error of the model. It is a measure of how much the predictions of the model deviate from the actual values.
 
-# Model {#sec-model}
+To put it another way, there exists a positive correlation between the educational attainment and societal status of parents with that of their children.
+However, when examining the specific factors of fathers' education and prestige score, it appears that the correlation with children's education is stronger than that of the mothers' education or prestige score.
 
-$$
-Pr(\theta | y) = \frac{Pr(y | \theta) Pr(\theta)}{Pr(y)}
-$$ {#eq-bayes}
-
-@eq-bayes seems useful, eh?
-
-Here's a dumb example of how to use some references: In paper we run our analysis in `R` [@citeR].
-We also use the `tidyverse` which was written by @thereferencecanbewhatever If we were interested in baseball data then @citeLahman could be useful.
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
+This suggests that fathers' educational background may have a particularly strong influence on their children's academic achievement and future opportunities.
+While the influence of mothers' education and social status is still significant, the impact of fathers' education appears to be even greater in shaping their children's educational outcomes.
+These findings highlight the importance of parental involvement and investment in their children's education, particularly that of the father's role in this regard.
 
@@
 
-## Weaknesses and next steps {#sec-weaknesses}
-*problems with prestige score:*
+## Weaknesses and next steps {#sec-weakness}
 
-Subjectivity: Even measures of prestige that attempt to capture objective indicators of social status, such as income or educational attainment, are still subject to subjective interpretation.
-People's perceptions of what constitutes a high-prestige occupation may vary based on their own cultural, social, or personal biases.
+As with any research, this study has its limitations and weaknesses.
+One significant weakness is the use of prestige scores as a variable to measure social status and occupational standing.
+While attempts were made to capture objective indicators of status, such as income and education, prestige scores are still subject to subjective interpretation.
+Cultural, social, or personal biases may influence people's perceptions of what constitutes a high-prestige occupation.
+Furthermore, measures of prestige are not fixed over time and may vary across different societies and historical periods.
+Heterogeneity within occupations can also lead to different levels of prestige among individuals.
+Additionally, some measures of prestige may overemphasize certain aspects of an occupation, such as income or education, while overlooking other critical factors like job security or working conditions [@prestige].
 
-Historical and cultural context: Measures of prestige are not fixed over time and may vary across different societies or historical periods.
-For example, certain occupations that were highly prestigious in the past, such as blacksmith or scribe, may be less highly regarded today.
+Another weakness of this study is missing data.
+While the decision to retain all missing values was made intentionally to avoid issues with representativeness, missing data can still affect the statistical power and accuracy of estimates.
+It can also reduce the representativeness of the sample, potentially limiting the generalizability of the findings.
+Additionally, missing data can make the analysis of the study more complex, requiring more sophisticated techniques that may increase the risk of errors [@kang].
 
-Heterogeneity within occupations: Different individuals within the same occupation may have different levels of prestige, depending on factors such as their level of experience, education, or specialization.
+Moreover, this study did not use hypothesis testing, which could potentially overlook some effects that might be present in the data.
+Therefore, the generalizability of the findings should be interpreted with caution.
+Future studies should aim to address these limitations and weaknesses to ensure more accurate and comprehensive research on the topic.
 
-Overemphasis on certain aspects of occupation: Some measures of prestige may place too much emphasis on certain aspects of occupation, such as income or educational requirements, while overlooking other important factors such as job security, working conditions, or level of autonomy.
+To shed light in future studies, there are some potential improvements.
+Firstly, it may be beneficial to incorporate additional measurements for social-economic statuses, such as the International Socio-Economic Index of Occupational Status (ISEI).
+This index is derived from the International Standard Classification of Occupations (ISCO) and comprises comparably coded data on education, occupation, and income from 73,901 full-time employed workers across 16 countries [@international].
 
-Lack of nuance: Prestige scores may provide a broad measure of social status and occupational major groups, but they may not capture the complexity and diversity of occupational experiences, especially for those in lower-prestige occupations.
-(https://www.researchgate.net/publication/236747304_A_Scale_of_Occupational_Prestige_in_Canada_Based_on_NOC_Major_Groups)
+Furthermore, new and accurate methodologies could be implemented for data analysis.
+It is evident that the sample sizes across the groups in @fig-birth are substantially different.
+Therefore, post-stratification may be a suitable technique to adjust for these discrepancies and verify the results.
+Moreover, in terms of the statistical model used in the study, a simple linear model was employed with only a few predictors.
+Future studies could build a more sophisticated model that accounts for potential interaction or quartic terms in the analysis to improve the understanding of the relationship under investigation.
 
-*problems with missing data:*
+In @fig-region, the results seem to be insignificant.
+One way to improve is to consider the measurement of the neighbourhood using living regions at age 16 may not provide a comprehensive representation of the impact of the neighbourhood on education.
+Other factors, such as the quality of schools and access to resources, may also play a significant role in educational outcomes but are not captured by this measure.
+Therefore, it is possible that the relationship between neighbourhood and education is more complex than what can be inferred from the living region at age 16 alone.
+While the distribution of education levels among neighbourhoods may appear similar based on @fig-region, it is still possible that there are significant differences in educational outcomes that are not apparent in the bimodal shape of the distribution.
+Further research using more detailed measures of neighbourhood characteristics and educational outcomes could provide a more nuanced understanding of the relationship between neighbourhood and education.
 
-Reduced statistical power is one such issue, which lowers the likelihood of rejecting the null hypothesis when it is false.
-Missing data can also cause bias in parameter estimation.
-The representativeness of the sample can also be reduced due to missing data.
-The analysis of the study can become more complex when missing data is present.
-These problems can threaten the validity of the trials and lead to invalid conclusions.
-(https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3668100/)
+While this data and study provide valuable insights into children's education and performance, it is important to acknowledge that some key factors have not been included.
+One such variable is the positive involvement of parents in their children's education.
+Research has consistently shown that parental involvement can have a significant impact on children's academic success, yet this aspect was not measured in the 2021 GSS [@parentinvolvement].
+Therefore, it is essential to include these factors in future studies to fully understand the complex interplay between different variables and their effects on children's education.
+By including factors like parental involvement, future studies could provide a more comprehensive understanding of the various factors that influence children's education and help to identify effective strategies for improving educational outcomes.
 

### 2023-03-27 scripts/02-data_cleaning.R
change in the title names and tests based on feedbacks

@@ -62,7 +62,5 @@ raw_data <- raw_data %>%
 cleaned_data<-raw_data
 
 #### Save data ####
-# [...UPDATE THIS...]
-# change cleaned_data to whatever name you end up with at the end of cleaning
-write_csv(cleaned_data, "inputs/data/GSS2021 2.csv")
+write_csv(cleaned_data, "inputs/data/GSS2021.csv")
 

### 2023-03-27 scripts/04-test_data.R

@@ -11,7 +11,7 @@
 #### Workspace setup ####
 library(tidyverse)
 cleaned_data <- read_csv(
-  file = here("inputs/data/GSS2021 2.csv"))
+  file = here("inputs/data/GSS2021.csv"))
 
 #### Test data ####
 cleaned_data$mapres10|> class() == "numeric"
@@ -45,3 +45,6 @@ cleaned_data$maeduc|> min(na.rm = T) >= 0
 cleaned_data$paeduc|>min(na.rm = T) >= 0
 
 cleaned_data$educ|>min(na.rm = T) >= 0
+
+#check none of the row was removed
+nrow(cleaned_data) == nrow(raw_data)

### 2023-03-27 scripts/00-simulate_data.R
changes in the scripts based on the feedback

@@ -1,10 +1,10 @@
 #### Preamble ####
-# Purpose: Data cleaning for GSS 2021
+# Purpose: Data Simulation for the paper of "The Relationship between Family Status and Academic Achievement in Children"
 # Author: Ruibo Sun
-# Data: 16 March 2023
+# Date: 16 March 2023
 # Contact: ruibo.sun@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: Download all the packges under the Workspace section
+# Pre-requisites: Download package of tidyverse under the Workspace section
 # Any other information needed? None
 
 

### 2023-03-27 scripts/01-download_data.R

@@ -1,26 +1,27 @@
 #### Preamble ####
-# Purpose: Downloads and saves the data from [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Data: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Download the data from GSS 2021
+# Author: Ruibo Sun
+# Date: 16 March 2023
+# Contact: ruibo.sun@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: Download all the packges under the Workspace section
+# Any other information needed? None
 
 
 #### Workspace setup ####
-library(opendatatoronto)
+library(haven)
 library(tidyverse)
-# [...UPDATE THIS...]
 
 #### Download data ####
-# [...ADD CODE HERE TO DOWNLOAD...]
+raw_data <- read_dta("inputs/data/GSS2021.dta")
 
+raw_data <-
+  labelled::to_factor(raw_data)
 
 
 #### Save data ####
 # [...UPDATE THIS...]
 # change the_raw_data to whatever name you assigned when you downloaded it.
-write_csv(the_raw_data, "inputs/data/raw_data.csv") 
+write_csv(raw_data, "inputs/data/raw_data.csv") 
 
          

### 2023-03-27 scripts/02-data_cleaning.R

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Data cleaning for GSS 2021
 # Author: Ruibo Sun
-# Data: 16 March 2023
+# Date: 16 March 2023
 # Contact: ruibo.sun@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: Download all the packges under the Workspace section
@@ -14,10 +14,7 @@ library(tidyverse)
 library(dplyr)
 
 #### Clean data ####
-raw_data <- read_dta("inputs/data/GSS2021.dta")
-
-raw_data <-
-  labelled::to_factor(raw_data)
+raw_data <- read_csv("inputs/data/raw_data.csv")
 
 raw_data <- 
   raw_data %>% 
@@ -62,5 +59,5 @@ raw_data <- raw_data %>%
 cleaned_data<-raw_data
 
 #### Save data ####
-write_csv(cleaned_data, "inputs/data/GSS2021.csv")
+write_csv(cleaned_data, "inputs/data/cleaned_GSS2021.csv")
 

### 2023-03-28 outputs/paper/paper.qmd
add the survey and its link! only discussion part left!

@@
 ---
-title: "The Relationship between Family Status and Academic Achievement in Children"
+title: "The Relationship between Family Status and Academic Achievement in Children in U.S. from 2021"
 author: 
@@
 date: "`r Sys.time()`"
-date-format: "12-03-2023"
+date-format: "D MMMM YYYY"
 abstract: "This paper explores the relationship between family status variables, such as parents' socio-economic status and educational level, and their impact on children's academic achievement and education level using the data from 2021 GSS.
@@
 cleaned_data <- read_csv(
-  file = here("inputs/data/GSS2021 2.csv"))
+  file = here("inputs/data/GSS2021.csv"))
 ```
@@
 
+# Appendix
+
+## Link to the Survey
+Please view the survey by this link [link](https://forms.office.com/Pages/ResponsePage.aspx?id=JsKqeAMvTUuQN7RtVsVSEIL6CTEUDb9LuJG06wrhLr9UMTNRNUNFS0tFT0hVTVM4UVFXWDdINFY5Sy4u).
+
+## Supplementary Survey
+1.What is your gender?
+
+- Male
+- Female
+- Non-binary
+- Woman
+- Man
+- Prefer not to say
+
+2.What is your race?
+
+- White
+- Black or African American
+- Hispanic or Latino
+- Asian
+- Native American or Alaska Native
+- Native Hawaiian or other Pacific Islander
+- Other
+
+3.Do you consider yourself as an immigrant?
+
+- Yes
+- No
+- Maybe
+
+4.US citizen
+
+- Yes
+- No
+
+5.What is the highest level of education you have achieved?
+
+- Less than high school
+- High school diploma or equivalent
+- Some college or associate degree
+- Bachelor's degree
+- Graduate degree or higher
+
+6.What type of school did you attend in your teenage?
+
+- Public school
+- Private school
+- Homeschool
+- Other
+
+7.High school GPA (if applicable)
+
+8.College/University GPA (if applicable)
+
+9.What is the highest level of education achieved by your mother and father?
+
+- Less than high school
+- High school diploma or equivalent
+- Some college or associate degree
+- Bachelor's degree
+- Graduate degree or higher
+
+10.What was your parents' occupation in your teenage?
+
+- Professional/managerial
+- Technical/sales
+- Administrative/clerical
+- Blue collar/other manual
+- Other
+
+11.What is the highest level of education completed by your mother?
+
+- Less than high school
+- High school graduate
+- Some college or technical school
+- Bachelor's degree
+- Graduate degree (Master's or Doctorate)
+
+12.What is the highest level of education completed by your father?
+
+- Less than high school
+- High school graduate
+- Some college or technical school
+- Bachelor's degree
+- Graduate degree (Master's or Doctorate)
+
+13.How involved are your parents in your education?
+
+- Not involved
+- Somewhat involved
+- Moderately involved
+- Highly involved
+
+14.What is your family's income level in your teenage?
+
+- Far above average
+- Above average
+- Average
+- Below average
+- Far below average
+
+15.Where did you live in your teenage years?
+
+- Rural area
+- Suburban area
+- Urban area
+
+16.How much pressure do you feel from your parents to excel academically?
+
+- None
+- A little
+- Moderate
+- High
+- Very high
+
 \newpage

### 2023-03-28 outputs/paper/paper.qmd
final draft of the paper

@@
 ---
-title: "The Relationship between Family Status and Academic Achievement in Children in U.S. from 2021"
+title: "The Relationship between Family Status and Academic Achievement in Children in US from 2021"
 author: 
@@
 
-# Results
+# Results {#sec-result}
 
@@
 
-## First discussion point {#sec-first-point}
+## Average years of schooling in the US and Outside of US
+Education has been a crucial driver of social and economic progress worldwide, and its expansion over the past two centuries has been a significant phenomenon.
+The global average year of schooling currently stands at around 8.7 years, according to data from World Economics [@world_economics].
+In this paper, by examining the average year of schooling in the United States, based on 2021 GSS data, it was found that the average year of schooling is 14 in @tbl-educ.
+This result is consistent with the findings from Our World in Data, which indicate that the average year of schooling in the US is also 14 years [@global_education].
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+Furthermore, the data in @tbl-degree reveals that a bachelor's degree and a graduate degree are the second and third most common educational categories in the US, respectively.
+This finding suggests that having an associate or junior college degree is less common compared to the bachelor's or graduate degree.
+Conversely, a smaller proportion of individuals have less than a high school education.
 
-## Second discussion point
+The Organisation for Economic Co-operation and Development (OECD) reports that the US is among the most well-educated countries globally, with 42% of 25-64 year-olds holding tertiary attainment, behind only Canada, Israel, Japan, and the Russian Federation [@glance].
+However, the US ranks 14th out of 37 OECD and G20 countries in the percentage of 25-34 year-olds with higher education, with 42% - above the OECD average but significantly behind Korea, which is the leader with 65%.
 
-## Third discussion point
+This data on education levels is not only important for understanding social and economic trends in the US, but it also has implications for policymaking and decision-making related to education.
+By accurately understanding the educational distribution of a given population, policymakers can make informed decisions about how to allocate resources and develop policies that address the needs and opportunities of different groups.
+
+## The impact of parental education level
+In the @sec-result, the @fig-parentseduc, @fig-paincome and @fig-maincome, they all show that there is a postive correlation between the education levels of mothers and fathers.
+If one parent has a high level of education, it is likely that the other parent also has a high level of education.
+
+The correlation between parents' education levels and their children's educational outcomes is well-established in the literature.
+Studies have consistently shown that parents' education is one of the strongest predictors of their children's academic achievement and educational attainment [@hout].
+The study by  [@chetty] used data from tax records and college enrollment records to analyse the relationship between parents' income, education, and social status, and their children's earnings and educational outcomes.
+The results showed that children from high-income families were more likely to attend college and earn higher incomes than children from low-income families.
+In addition, the study found that parental education was a key factor in determining children's educational and economic outcomes.
+Children with highly educated parents were more likely to attend college and earn higher incomes than children with less educated parents.
+
+Furthermore, the analysis of the statistical model from the @tbl-lmeduc shows that parental education and prestige scores are positively correlated with their children's academic achievement and societal status.
+The coefficients for the independent variables father’s and mother’s education in years indicate that an increase in fathers' education has a greater association with their children's education than that of mothers.
+This suggests that the father's role in education and investment in their children's future may be particularly important.
+However, the influence of both parents is still significant.
+The R-squared value suggests that the independent variables explain only a small portion of the variation in the dependent variable, indicating that other factors beyond parental education and prestige score also play a role.
+Nonetheless, these findings highlight the importance of parental involvement in their children's education, with a particular emphasis on the role of fathers.
+
+The relationship between parental education and student achievement has been the focus of many studies, with varying results.
+Some study has found that maternal educational attainment had an impact on student achievement, which contradicts the finding in this paper [@1993].
+The present study showed that paternal support of education has a significant impact on academic achievement, while maternal education level did not show a significant impact.
+However, the study of @2000 showed that families with active fathers fostered maternal involvement in the household, which resulted in a support system that fosters educational attainment and positive long-term contributions.
+
+Despite these discrepancies, the findings in this paper concur with prior literature that emphasizes the importance of parental education levels and their impact on college student achievement [@ortiz].
+This study focuses on showing that the fathers' education level has a significant relationship with childrens’ year of schooling, while no statistically significant relationship was observed with the mothers' education level.
+
+These findings highlight the importance of parental involvement and support in their children's education, with a particular emphasis on the role of fathers.
+The results suggest that fathers' education levels may have a stronger impact on their children's academic achievement than that of mothers.
+However, it is important to note that other factors, such as family income and race, may also play a role in student achievement.
+Therefore, further research is necessary to fully understand the complex relationship between these variables and their impact on academic achievement, which will be carefully discussed in @sec-weakness.
 
@@
 
+### Weaknesses and Limitations
 As with any research, this study has its limitations and weaknesses.
@@
 
+### Next Steps
 To shed light in future studies, there are some potential improvements.
@@
 
+Lastly, the study's data was limited to individuals who responded to the 2021 GSS survey in the US, which may not be representative of the entire population.
+Respondents who chose to participate in the survey may have different characteristics than non-respondents, which could affect the generalizability of the study's findings to the wider population.
+
 # Appendix
@@
 ## Link to the Survey
-Please view the survey by this link [link](https://forms.office.com/Pages/ResponsePage.aspx?id=JsKqeAMvTUuQN7RtVsVSEIL6CTEUDb9LuJG06wrhLr9UMTNRNUNFS0tFT0hVTVM4UVFXWDdINFY5Sy4u).
+Please view the survey by this [link](https://forms.office.com/Pages/ResponsePage.aspx?id=JsKqeAMvTUuQN7RtVsVSEIL6CTEUDb9LuJG06wrhLr9UMTNRNUNFS0tFT0hVTVM4UVFXWDdINFY5Sy4u).
 

### 2023-03-28 scripts/04-test_data.R
change "date"

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Data cleaning for GSS 2021
 # Author: Ruibo Sun
-# Data: 16 March 2023
+# Date: 16 March 2023
 # Contact: ruibo.sun@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: Download all the packges under the Workspace section

### 2023-04-03 outputs/paper/paper.qmd
changes based on feedback

@@
 date-format: "D MMMM YYYY"
-abstract: "This paper explores the relationship between family status variables, such as parents' socio-economic status and educational level, and their impact on children's academic achievement and education level using the data from 2021 GSS.
+abstract: "This paper explores the relationship between family status variables, such as parents' socio-economic status and educational level, and their impact on children's academic achievement and education level using the data from 2021 General Social Survey (GSS).
 Previous studies have suggested that parents' education can positively influence their children's academic performance due to increased access to resources and greater involvement in their education.
@@
 They also tend to hold higher expectations for their children's academic performance [@davis_2014].
+Ultimately, these may affect the education level of the children.
 
@@
 
+This following findings are from the 2021 General Social Survey (GSS) data.
+
 ```{r}
@@
   # geom_histogram(bins = 15, fill = "lightblue", color = "white") +
-  labs(x = "Education (in Years)", y = "Desensity") +
+  labs(x = "Education (in Years)", y = "Density") +
   scale_x_continuous(breaks = seq(0, 20, by = 1)) +
@@
 In @fig-educ, this histogram shows an overall distribution of the respondents' education in years.
-It is clear that the distribution is right-skewed, which indicates that there are more respondents with lower levels of education than with higher levels of education.
+It is clear that the distribution is right-skewed, which indicates that there are more respondents with higher levels of education.
 In other words, the bulk of the respondents are clustered towards the left side of the histogram (lower levels of education), and there are relatively fewer respondents towards the right side of the histogram (higher levels of education).
@@
   labs(x = "Education (in Years)",
-       y = "Mother's Educaion (in Years)",
+         y = "Mother's Education (in Years)",
        color = "Family Income (at Age of 16)") +
@@
   labs(x = "Education (in Years)",
-       y = "Father's Educaion (in Years)",
+       y = "Father's Education (in Years)",
   color = "Family Income (at Age of 16)") +

### 2023-04-05 outputs/paper/paper.qmd
intial draft only with few changes

@@
 ---
-title: "The Relationship between Family Status and Academic Achievement in Children in US from 2021"
+title: "Exploring the Link Between Family Background and Children's Academic Achievement in the United States (2021)"
 author: 
   - Ruibo Sun
-thanks: "https://github.com/ruibosun/how-parents-affect-childrens-education"
+thanks: "Code and Data: https://github.com/ruibosun/revised-how-parents-affect-childrens-education"
 date: "`r Sys.time()`"
@@
 cleaned_data <- read_csv(
-  file = here("inputs/data/GSS2021.csv"))
+  file = here("inputs/data/cleaned_GSS2021.csv"))
 ```
@@
 Instead, socio-economic status and parents' education are part of a larger set of psychological and sociological variables that can impact children's educational outcomes [@khan].
+
+In my previous paper, there are several findings.
+The average years of schooling in the US is 14, which is higher than the global average of 8.7 years.
+A bachelor's degree and a graduate degree are more common than an associate or junior college degree, and the US ranks behind several countries in terms of higher education attainment for 25-34 year-olds.
+Parental education levels have a positive correlation, and there is a strong link between parents' education levels and their children's academic achievement and societal status.
+Fathers' education levels appear to have a stronger impact than mothers' on children's academic achievement, but other factors may also play a role.
+Further research is needed to fully understand the relationship between these variables.
+Policymakers can use this information to make informed decisions about resource allocation and policy
+development.
+Here is a [link](https://github.com/ruibosun/how-parents-affect-childrens-education/blob/main/outputs/paper/paper.pdf) to view the entire paper.However, the prior study did not appropriately handle the issue of missing values (NAs).
+Consequently, this new paper aims to conduct a more rigorous analysis by appropriately addressing the NAs.
+Moreover, it will address the limitations of the previous study to produce more accurate and representative results.
 
@@
 - **educ**: A numeric variable indicating the number of years of education completed by the respondent.
-- **mawrkslf**: A categorical variable indicating whether the respondent's mother was employed for wages or salary.
-- **pawrkslf**: A categorical variable indicating whether the respondent's father was employed for wages or salary.
 - **degree**: A categorical variable indicating the highest degree earned by the respondent.
@@
 
-# Results {#sec-result}
+# Results* {#sec-result}
 
@@
 
-# Model {#sec-model}
+# Model* {#sec-model}
 
@@
 # Fit the model
-model <- lm(educ ~ mapres10 + papres10 + paeduc + maeduc, data = cleaned_data)
+model <- lm(educ ~ mapres10 + papres10 + exp(paeduc) + exp(maeduc), data = cleaned_data)
 
@@
 
-# Discussion
+# Discussion*
 
@@
 
-## Weaknesses and next steps {#sec-weakness}
+## Weaknesses and next steps* {#sec-weakness}
 

### 2023-04-05 scripts/02-data_cleaning.R

@@ -4,7 +4,7 @@
 # Date: 16 March 2023
 # Contact: ruibo.sun@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: Download all the packges under the Workspace section
+# Pre-requisites: Download all the packges under the Workspace section and read the 2021 General Social Survey data.
 # Any other information needed? None
 
 
@@ -12,52 +12,66 @@
 library(haven)
 library(tidyverse)
 library(dplyr)
+library(mice)
 
 #### Clean data ####
 raw_data <- read_csv("inputs/data/raw_data.csv")
 
 raw_data <- 
   raw_data %>% 
-  dplyr::select(mapres10,papres10,paeduc,maeduc,educ,mawrkslf,pawrkslf,degree,born,sex,reg16,incom16)
+  dplyr::select(mapres10,papres10,paeduc,maeduc,educ,degree,born,reg16,incom16)
 
-#change class
-raw_data$mapres10 <- as.character(raw_data$mapres10)
-raw_data$papres10 <- as.character(raw_data$papres10)
-raw_data$maeduc <- as.character(raw_data$maeduc)
-raw_data$paeduc <- as.character(raw_data$paeduc)
-raw_data$educ <- as.character(raw_data$educ)
 
-# Replace "no formal schooling" with 0 in the specified column
-raw_data$educ[raw_data$educ== "no formal schooling"] <- 0
-raw_data$maeduc[raw_data$maeduc== "no formal schooling"] <- 0
-raw_data$paeduc[raw_data$paeduc== "no formal schooling"] <- 0
+# Columns to process
+cols_to_process <- c("mapres10", "papres10", "maeduc", "paeduc", "educ")
 
-# Change to numeric
-raw_data$mapres10 <- as.numeric(raw_data$mapres10)
-raw_data$papres10 <- as.numeric(raw_data$papres10)
-raw_data$maeduc <- as.numeric(raw_data$maeduc)
-raw_data$paeduc <- as.numeric(raw_data$paeduc)
-raw_data$educ <- as.numeric(raw_data$educ)
+# Loop through columns and process
+for (col in cols_to_process) {
+  # Change to character
+  raw_data[[col]] <- as.character(raw_data[[col]])
+  
+  # Replace "no formal schooling" with 0
+  raw_data[[col]][raw_data[[col]] == "no formal schooling"] <- 0
+  
+  # Change to numeric
[…]

### 2023-04-05 scripts/04-test_data.R

@@ -46,5 +46,5 @@ cleaned_data$paeduc|>min(na.rm = T) >= 0
 
 cleaned_data$educ|>min(na.rm = T) >= 0
 
-#check none of the row was removed
-nrow(cleaned_data) == nrow(raw_data)
+# check the NAs that are numerical are gone
+apply(X = is.na(cleaned_data), MARGIN = 2, FUN = sum)==c(rep(0,9))

### 2023-04-11 outputs/paper/paper.qmd
add the model for prestige score and add text to data cleaning

@@
 development.
-Here is a [link](https://github.com/ruibosun/how-parents-affect-childrens-education/blob/main/outputs/paper/paper.pdf) to view the entire paper.However, the prior study did not appropriately handle the issue of missing values (NAs).
+Here is a [link](https://github.com/ruibosun/how-parents-affect-childrens-education/blob/main/outputs/paper/paper.pdf) to view the entire paper.
+However, the prior study did not appropriately handle the issue of missing values (NAs).
 Consequently, this new paper aims to conduct a more rigorous analysis by appropriately addressing the NAs.
@@
 The package inside tidyverse helps to create the plots in ggplot2 [@tidyverse].
+The missing values were imputed using the mice package [@mice].
 This paper is knitted as a PDF file by the packages of R markdown [@RMarkdown] and formatted using patchwork [@patchwork].
@@
 The GSS employs various methods to adjust for non-response and ensure the representativeness of the sample.
-One strength of the GSS is its long history and large sample size, but a potential weakness is a reliance on self-reported data, which can be subject to response bias.
 The questionnaire includes both closed-ended and open-ended questions, but there may be concerns about biased or leading questions.
@@
 
+$$
+Prestige_{ij}=\mu+\alpha_i+\beta_{j}+\epsilon_{ij}
+$$ {#eq-modpr}
+
 This paper will test and explore how parents’ social-economics status can affect children’s education.
 The data that will be used in this paper comes from the US General Social Survey from the National Opinion Research Center at the University of Chicago.
+
 The factor of socio-economics status is measured using the occupation prestige score.
 It is a measure used in social science research to assess the level of social status or prestige associated with a particular occupation or profession, which was developed by the National Opinion Research Center at the University of Chicago.
-As part of the GSS, respondents are asked to rate the prestige or social standing of various occupations on a scale of 1 to 90, with higher scores indicating greater prestige.
-[@smith] Another variable that has been used is the education level, which is a number of variables to indicate the number of years the respondents have spent in school and college.
+As part of the GSS, respondents are asked to rate the prestige or social standing of various occupations on a scale of 1 to 100, with higher scores indicating greater prestige.
+The ideal dataset would have 90,090 ratings from 1,001 raters, but due to various issues, only 82,800 ratings from 946 raters were used.
+Invalid or inconsistent ratings, such as cases with reversed codes, too few ratings, equal scores for all occupations, or a low standard deviation of ratings, were excluded to ensure the reliability of the occupational scores.[@smith]
+
+Next,the prestige score is obtained by the @eq-modpr.
+$Prestige_{ij}$ is the prestige score for occupational title i and rater j.
+The $\alpha_i$ represents the occupational differences of interest, while the $\beta_j$ represents differences among the raters that need to be controlled for in estimating the $\alpha_i$.
+The $\mu$ is the overall mean prestige score, and $\epsilon_{ij}$ represents the random error term.
+The range of the prestige scores is from 0 to 100, with higher scores indicating a better or more prestigious occupation[@smith].
+
+Another variable that has been used is the education level, which is a number of variables to indicate the number of years the respondents have spent in school and college.
 Other variables will be discussed in @sec-data.
@@
 For simplicity purposes, only variables that are closely related to the topics are selected for further analysis.
-These variables are mapres10, papres10, paeduc, maeduc, educ, mawrkslf, pawrkslf, degree, born, sex, reg16, and incom16.
+These variables are **mapres10**, **papres10**, **paeduc**, **maeduc**, **educ**, **degree**, **born**, **sex**, **reg16**, and **incom16**.
 
@@
 
-None of the missing values (NA) is removed in this paper.
-The reason is that removing the missing values in every column would result in a huge loss in the dataset.
-To maintain the representativeness of the data, missing values are automatically filtered by R, rather than removed manually.
-The possible limitation and effect of the missing values will be discussed in the @sec-weakness.
+As mentioned before, this paper is an revised version of [the previous paper](https://github.com/ruibosun/how-parents-affect-childrens-education/blob/main/outputs/paper/paper.pdf) which has some limitations and weaknesses.
+In order to improve the representiveness of the data, this paper use the multiple-imputation and dropping methods [@gelman].
+There are two types of missing values: numercial and categorical missing values.
+
+For numerical variables (**paeduc**, **maeduc**, **educ**, **mapres10** and **papres10**), multiple imputation techniques are employed to estimate and fill in the missing data points.
+This method generates a more accurate dataset without the missing values.
+Following the imputation, the resulting dataset is transformed into a new one with no NAs for further analysis.
+
+Next, any categorical variables with missing values are addressed, such as **degree**, **sex**, **born**, **reg16** and **incom16**.
+In this case, all these variable with NAs are removed from the data, ensuring that the analysis is only performed on complete cases.
+This helps to minimize any biases that might arise due to incomplete data.
 

### 2023-04-12 outputs/paper/paper.qmd
almost finish the model, lefe with discussion

@@
 ---
-title: "Exploring the Link Between Family Background and Children's Academic Achievement in the United States (2021)"
+title: "Exploring the Relation between Family Background and Children's Education Level in the United States (2021)"
 author: 
@@
 Previous studies have suggested that parents' education can positively influence their children's academic performance due to increased access to resources and greater involvement in their education.
-However, recent research has demonstrated that this relationship is complex and can be influenced by a variety of psychological and sociological factors.
-The findings suggest that parents' education, family income, and prestige score all have a positive correlation with children's education.
+However, recent research has demonstrated that this relationship is complex and can be influenced by a variety of factors.
+The findings in this paper suggest that parents' education, family income, and prestige score all affect the children's education level.
 Further research is needed to fully understand the complexities of this relationship and use the results to improve educational policies and programs."
@@
 The data that will be used in this paper comes from the US General Social Survey from the National Opinion Research Center at the University of Chicago.
+In this study, the "children" are the respondents and "parent" are the parents of the respondents.
 
@@
 
-## Data cleaning
+## Data cleaning {#sec-cleaning}
 
 For simplicity purposes, only variables that are closely related to the topics are selected for further analysis.
-These variables are **mapres10**, **papres10**, **paeduc**, **maeduc**, **educ**, **degree**, **born**, **sex**, **reg16**, and **incom16**.
+These variables are **mapres10**, **papres10**, **paeduc**, **maeduc**, **educ**, **degree**, **born**, **reg16**, and **incom16**.
 
@@
 - **born**: A categorical variable indicating the region of the US where the respondent was born.
-- **sex**: A categorical variable indicating the respondent's sex.
 - **reg16**: A categorical variable indicating the region of the US where the respondent lived at age of 16.
@@
 
-Next, any categorical variables with missing values are addressed, such as **degree**, **sex**, **born**, **reg16** and **incom16**.
+Next, any categorical variables with missing values are addressed, such as **degree**, **born**, **reg16** and **incom16**.
 In this case, all these variable with NAs are removed from the data, ensuring that the analysis is only performed on complete cases.
@@
 
-# Results* {#sec-result}
+The number of oberservation before data cleaning is 4032. This number decrease to 3794 because some rows with missing values. are dropped from the data.
+
+In addtion, the variable of **degree** is grouped into two levels.
+For respondents who have obtained at least a degree from an associate/junior college, their levels are denoted as "1", indicating they have complete a post-secondary education.
+For respondents who have attained only a high school diploma or possess a lower level of education, their levels are designated as "0".
+
+# Results {#sec-result}
 
@@
   mutate(percent=percent(x=n/sum(n)))
-tb1<-tb1%>%arrange(factor(tb1$degree,c("graduate", "bachelor's", "associate/junior college", "high school", "less than high school", NA)))
+tb1<-tb1%>%arrange(factor(tb1$degree,c("graduate", "bachelor's", "associate/junior college", "high school", "less than high school")))
 colnames(tb1) <- c("Degree", "Total", "Proportion")
@@
 ```
-The respondents in this study are the children and their parents' education has also been collected in the GSS.
-As for the response variables, children’s education level, there are two ways to measure.
-One is the degree, and the other one is the number of years in school.
-In @tbl-degree, it shows the distribution of education in terms of degrees.
-The data reveals that the most common degree category among the group is high school, with 1597 individuals (39.6%) having this level of education.
-Following this, the next most common category is a bachelor's degree, with 1036 individuals (25.7%) having completed this level of education.
-The data also indicates that a graduate degree is the third most common category, with 760 individuals (18.8%) having this level of education.
-Associate/junior college degree is less common than a bachelor's or graduate degree, with 370 individuals (9.2%) having completed this level of education.
-However, a smaller proportion of individuals (6.1%) have less than a high school education.
-Additionally, there are also 23 individuals (0.6%) whose educational degree information is not available (NA).
-Overall, the data provide insights into the educational distribution of the given group and may be useful for drawing conclusions and making informed decisions related to education.
+According to the @tbl-degree, the distribution of education levels in terms of degrees shows that the most prevalent category is high school education, with 1507 individuals (39.7%) falling into this group.
+Next in prevalence is a bachelor's degree, held by 1010 individuals (26.6%).
+A graduate degree is the third most common category, with 731 individuals (19.3%) having attained this level of education.
+In contrast, the associate/junior college degree is less frequent, with 362 individuals (9.5%) completing this level of education.
+The smallest proportion, 4.8%, have less than a high school education.
 
@@
   geom_histogram(aes(y = ..density..), binwidth = 2, fill = "white") +
-  geom_density(alpha = .2, fill = "lightblue") +
+  geom_density(alpha = .2, fill = "green") +
 
@@
   labs(x = "Education (in Years)", y = "Density") +
-  scale_x_continuous(breaks = seq(0, 20, by = 1)) +
   theme_bw()
@@
                               Max = max(cleaned_data$educ,na.rm = T),
-                              Mean = mean(cleaned_data$educ,na.rm=T))
+                              Mean = round(mean(cleaned_data$educ,na.rm=T),digits=2))
 
@@
 To see a more accurate numerical summary, @tbl-educ is created.
-In the dataset labelled @sec-data, the category denoting "no formal schooling" has been redefined as an integer value of 0.
+In the dataset labelled @sec-data, the category denoting "no formal schooling" has been redefined as an integer value, 0.
 According to the information presented in the @tbl-educ table, the range of years of schooling reported by respondents spans from 0 to 20 years.
-The median number of years of schooling reported is 15, and the mean is 14.76904.
-This mean value roughly corresponds to three years in college.
+The median number of years of schooling reported is 15, and the mean is 14.82 This mean value roughly corresponds to three years in college.
 Taken together, these findings suggest that the respondents surveyed in the 2021 GSS have generally completed their high school education and are highly educated.
@@
 The distribution of father’s prestige score and the mother’s prestige score have quite different distributions.
+Specifically, the distribution of mother's prestige score is bimodal and the distribution of father's prestige score is unimodal.
 
@@
 ggplot(cleaned_data, aes(x = maeduc)) +
-geom_histogram(aes(y = ..density.., fill = "Mother's Education"), alpha = 1, binwidth = 0.5, color = "black") +
-geom_histogram(aes(x = paeduc, y = ..density.., fill = "Father's Education"), alpha = 0.8, binwidth = 0.5, color = "black") +
+  geom_histogram(aes(y = ..density.., fill = "Mother's Education"),
+                 alpha = 1, binwidth = 0.5, color ="black") +geom_histogram(aes(x = paeduc, y = ..density.., fill = "Father's Education"),alpha = 0.8, binwidth = 0.5, color = "black") +
 scale_fill_manual(values = c("Mother's Education" = "darkgreen", "Father's Education" = "lightgreen"), name = "Variables") +
@@
 ggplot(cleaned_data, aes(x = educ, fill = factor(born))) +
-  geom_histogram(position = "dodge", binwidth = 1.5, alpha = 0.5, color = "black")+
+  geom_histogram(aes(y = ..density..), position = "dodge", binwidth = 1.5, alpha = 0.5, color = "black")+
    labs(x = "Education (in Years)",
-       y = "Count",
+       y = "Density",
        fill = "Birthplace of the Respondent (if in US)")+
   theme_bw()
+
+# ggplot(cleaned_data, aes(x = educ)) +
+# geom_histogram(aes(y = ..density.., fill = "Mother's Education"), alpha = 1, binwidth = 0.5, color = "black") +
+# geom_histogram(aes(x = paeduc, y = ..density.., fill = "Father's Education"), alpha = 0.8, binwidth = 0.5, color = "black") +
+# scale_fill_manual(values = c("Mother's Education" = "darkgreen", "Father's Education" = "lightgreen"), name = "Variables") +
+# labs(x = "Education (in Years)",
+# y = "Density") +
+# theme_bw()
 ```
@@
 However, the bar for respondents who were born outside of the US is much higher than the other groups.
-This is because 87.2% of the respondents were born in the US, and only 11.0% answered “no”.
-Additionally, there are 72 NA, which take a proportion of 1.8%.
-Given the huge differences in the sample sizes between groups, it is hard to draw a conclusion of whether birthplace may affect the education level.
+This is because 90% of the respondents were born in the US, and only 10.0% answered “No”.
+This is because respondents who did not answer this question or answered NA are removed during the data cleaning.
+Given the huge differences in the sample sizes between groups, it is hard to draw a general and representative conclusion of whether birthplace may affect the education level.
 
@@
 ```
+
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+#| label: tbl-regandeduc
+#| tbl-cap: The proportion and mean of education for different regions
+tb_reg <- cleaned_data %>% 
+  count(reg16) %>% 
+  mutate(percent=percent(x=n/sum(n)))
+
+tb_reg_avg<-cleaned_data %>%
+  group_by(reg16) %>%
+  summarise(mean_educ = round(mean(educ),digits=2))
+tb_reg<-tb_reg%>%mutate("Mean of Education"=tb_reg_avg$mean_educ)
+colnames(tb_reg)<-c("Region","Count","Percent","Mean of Education")
+tb_reg %>% kable()
+```
+The @tb-regandeduc showcases the average number of years of education for individuals in different living regions of the United States when the respondent's was 16.
+The highest average education level is found in the Middle Atlantic region, where the mean is 15.13 years.
+The East North Central region follows closely, with an average of 14.98 years of education.
+In the Pacific and New England regions, the mean years of education are relatively similar, with 14.94 and 14.85 years, respectively.
+Those regions are above the overall average, which is 14.82 from @tb_educ.
+On the other hand, the South Atlantic and West North Central regions also display comparable average education levels of 14.75 and 14.73 years, respectively.
+The Mountain region exhibits a slightly lower mean of 14.69 years.
+Furthermore, the East South Atlantic and West South Central regions have the lowest average education levels, with 14.10 and 14.39 years, respectively.
 
@@
 ggplot(cleaned_data, aes(x = educ,fill = factor(reg16))) +
-  geom_histogram(binwidth = 2, color = "black") +
-  facet_wrap(~ reg16, ncol = 2)+
+  geom_histogram(aes(y = ..density..),binwidth = 2, color = "black") +
+  facet_wrap(~ reg16, ncol = 3)+
   labs(x = "Education (in Years)",
-       y = "Count")+
+       y = "Density")+
   theme_bw()+
@@
 In the 2021 GSS, the neighbourhood variable is being measured by the living region at the age of 16.
-The distribution for all regions, as shown in @fig-region, follows a similar bimodal shape, with the exception of the NA group.
+The distribution for all regions, as shown in @fig-region, follows a similar bimodal shape.
 This could indicate that the region of living has a minimal or no impact on children's education levels.
 However, it is important to note that the measurement of the neighborhood through living regions at age 16 may not fully capture the impact of the neighborhood on education, which will be discussed in the @sec-weakness.
-
-```{r,echo=FALSE}
-tb4 <- cleaned_data %>% 
-  count(reg16) %>% 
-  mutate(percent=percent(x=n/sum(n)))
-```
 
@@
 In @fig-maincome, the vertical axis represents the mother’s education and the horizontal axis represents the children’s education.
-The respondents were asked to rate their family income at age of 16 in the following category: (1) far above average, (2) above average, (3) average, (4) below average and (5) far below average.
-Due to missing values, the sixth group is NA, which is coloured grey.
+The respondents were asked to rate their family income at age of 16 in the following category: (1) **far above average**, (2) **above average**, (3) **average**, (4) **below average** and (5) **far below average**.
 Overall, mothers’ education and children’s education are positively correlated despite the level of income.
-To be more specific, for respondents whose family income was far above average, the slope seems to be steeper than the other groups, which indicates that for those with higher family income, there may be a stronger relationship between the education of the mother and the education of the children.
+To be more specific, for respondents whose family income was **far above average**, the slope seems to be steeper than the other groups, which indicates that for those with higher family income, there may be a stronger relationship between the education of the mother and the education of the children.
 
@@
 Once again, there is a positive correlation between the education level of fathers and their children's education, regardless of family income.
-However, for respondents who reported a "far above average" family income, the relationship between the father's education and children's education appears to be stronger than for other income groups.
+However, for respondents who reported a **far above average** family income, the relationship between the father's education and children's education appears to be stronger than for other income groups.
 This suggests that higher family income may be associated with a more pronounced link between paternal and children's education levels.
@@
 $$
-\hat{Y}=\hat{\beta_0}+\hat{\beta_1}X_{mapres10}+\hat{\beta_2}X_{papres10}+\hat{\beta_3}X_{paeduc}+\hat{\beta_4}X_{maeduc}
+\log(\frac{\pi}{1-\pi})=\hat{\beta_0}+\hat{\beta_1}X_{mapres10}+\hat{\beta_2}X_{papres10}+\hat{\beta_3}X_{paeduc}+\hat{\beta_4}X_{maeduc}
 $$ {#eq-model}
@@
 #| label: tbl-lmeduc
-#| tbl-cap: Linear Model and its Summary Statistics
+#| tbl-cap: Logistic Model and its Summary Statistics
 
 # Fit the model
-model <- lm(educ ~ mapres10 + papres10 + exp(paeduc) + exp(maeduc), data = cleaned_data)
+model <- glm(high_school ~ mapres10 + papres10 + paeduc + maeduc + incom16 + reg16, family="binomial", data = cleaned_data)
 
@@
 ```
-The @tbl-lmeduc is built based on the @eq-model:
+The @tbl-lmeduc is built based on the @eq-model. In this logistic model, it is used to predict the probability of a child attains a post-secondary degree based on these predictors.
+*missing the bullet point to explain each exp(coef)*
 
-- The intercept is 11.32, which means that when all the independent variables are zero, the dependent variable has a value of 11.32.
-- The coefficients for the independent variables mapres10 and papres10 are both 0.01, which means that a one-unit increase in either of these variables is associated with an increase of 0.01 in the dependent variable.
-- The coefficients for the independent variables paeduc and maeduc are 0.18 and 0.05, respectively.
-This means that a one-unit increase in paeduc is associated with an increase of 0.18 in the dependent variable, and a one-unit increase in maeduc is associated with an increase of 0.05 in the dependent variable.
-- The number of observations used in the model is 1980.
-The R-squared value is 0.124, which means that the independent variables explain 12.4% of the variation in the dependent variable.
-- The adjusted R-squared value is 0.122.
-- The AIC and BIC values are measures of the goodness of fit of the model. Lower values indicate a better fit.
-- The log-likelihood is a measure of how well the model fits the data.
-- The RMSE value is 2.44, which is the root mean squared error of the model. It is a measure of how much the predictions of the model deviate from the actual values.
+From @tbl-lmeduc,the positive coefficients for both mother's and father's prestige scores (**mapres10** and **papres10**) suggest a slight positive relationship with children's education levels.
+Moreover, the positive coefficients for parents' education (**paeduc** and **maeduc**) have a greater positive association with children's education level.
 
-To put it another way, there exists a positive correlation between the educational attainment and societal status of parents with that of their children.
-However, when examining the specific factors of fathers' education and prestige score, it appears that the correlation with children's education is stronger than that of the mothers' education or prestige score.
+The coefficients for the income level at age of 16 (**incom16**) indicate that higher-income levels are generally associated with lower education levels for children.
+Particularly, the "far above average" and "far below average" income categories show a significant negative relationship with children's education levels.
 
-This suggests that fathers' educational background may have a particularly strong influence on their children's academic achievement and future opportunities.
-While the influence of mothers' education and social status is still significant, the impact of fathers' education appears to be even greater in shaping their children's educational outcomes.
-These findings highlight the importance of parental involvement and investment in their children's education, particularly that of the father's role in this regard.
+Regarding the living regional differences at age of 16 (**reg16**), the results indicate that children from the east south atlantic region may have lower education levels compared to those from other regions.
+The positive coefficients for middle atlantic and west north central regions suggest a positive association with children's education levels, although the results are not statistically significant.
 
@@
 
-## Weaknesses and next steps* {#sec-weakness}
+## The Effect of Living Region during the Adolescent Period
+referring results from the model about how regions affect the education level
+
+## Weaknesses and next steps {#sec-weakness}
 
@@
 Additionally, some measures of prestige may overemphasize certain aspects of an occupation, such as income or education, while overlooking other critical factors like job security or working conditions [@prestige].
-
-Another weakness of this study is missing data.
-While the decision to retain all missing values was made intentionally to avoid issues with representativeness, missing data can still affect the statistical power and accuracy of estimates.
-It can also reduce the representativeness of the sample, potentially limiting the generalizability of the findings.
-Additionally, missing data can make the analysis of the study more complex, requiring more sophisticated techniques that may increase the risk of errors [@kang].
 
@@
 
-In @fig-region, the results seem to be insignificant.
+In the @fig-region, the results seem to be insignificant.
 One way to improve is to consider the measurement of the neighbourhood using living regions at age 16 may not provide a comprehensive representation of the impact of the neighbourhood on education.
@@
 Further research using more detailed measures of neighbourhood characteristics and educational outcomes could provide a more nuanced understanding of the relationship between neighbourhood and education.
+
+In the @tbl-lmeduc, a logistic model is employed with the degree as the response variable.
+However, using a linear regression model with years of education as the response variable may provide a more detailed examination of the factors involved.
+A linear regression can help analyze specific relationships, such as how years of education change with a five-unit increase in parents' prestige scores.
+This study opted not to use the linear model due to violations of underlying assumptions.
+Future research may improve upon this by constructing a more sophisticated linear regression model to elucidate the relationship between these variables in a more precise manner.
 

### 2023-04-12 scripts/02-data_cleaning.R

@@ -71,6 +71,10 @@ cleaned_data <- cleaned_data %>%
     mapres10 >= mema ~ "Above and Equal to Median"
   ))
 
+## for the model
+cleaned_data <- cleaned_data %>%
+  mutate(high_school = ifelse(degree %in% c("graduate", "bachelor's", "associate/junior college"),1,0))
+
 #### Save data ####
 write_csv(cleaned_data, "inputs/data/cleaned_GSS2021.csv")
 

### 2023-04-12 scripts/00-simulate_data.R
Minor changes

@@ -1,10 +1,10 @@
 #### Preamble ####
 # Purpose: Data Simulation for the paper of "The Relationship between Family Status and Academic Achievement in Children"
 # Author: Ruibo Sun
-# Date: 16 March 2023
+# Date: 12 April 2023
 # Contact: ruibo.sun@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: Download package of tidyverse under the Workspace section
+# Pre-requisites: None
 # Any other information needed? None
 
 
@@ -13,41 +13,65 @@ library(tidyverse)
 
 
 #### Simulate data ####
-set.seed(123) # set seed for reproducibility
+set.seed(123)
+n <- 100
 
-# generate 100 observations
-n <- 100 
+# Generate individual variables
+parent_edu <-
+  sample(
+    c(
+      "less than high school",
+      "high school",
+      "some college",
+      "bachelor's degree",
+      "graduate degree"
+    ),
+    n,
+    replace = TRUE
+  )
 
-# generate parents' education level, with higher levels of education having a greater probability of occurrence
-parent_edu <- sample(c("less than high school", "high school", "some college", "bachelor's degree", "graduate degree"), 
-                     n, 
-                     replace = TRUE)
+parent_higher_edu <-
+  ifelse(parent_edu %in% c("bachelor's degree", "graduate degree"), 1, 0)
 
-# generate a binary variable indicating whether parents have higher education or not
-parent_higher_edu <- ifelse(parent_edu %in% c("bachelor's degree", "graduate degree"), 1, 0)
+child_edu <-
+  sample(
+    c(
+      "less than high school",
+      "high school",
+      "some college",
+      "bachelor's degree",
+      "graduate degree"
+    ),
+    n,
+    replace = TRUE
+  )
 
-# generate children's education, with higher probability of occurrence if parents have higher education
-child_edu <- sample(c("less than high school", "high school", "some college", "bachelor's degree", "graduate degree"), 
-                    n, 
-                    replace = TRUE)
+child_higher_edu <-
+  ifelse(child_edu %in% c("bachelor's degree", "graduate degree"), 1, 0)
 
-# generate a binary variable indicating whether children have higher education or not
-child_higher_edu <- ifelse(child_edu %in% c("bachelor's degree", "graduate degree"), 1, 0)
+region <-
+  sample(c("North", "South", "East", "West"), n, replace = TRUE)
 
[…]

### 2023-04-16 outputs/paper/paper.qmd

@@
 ---
-title: "Exploring the Relation between Family Background and Children's Education Level in the United States (2021)"
+title: "Exploring the Relationship between Family Background and Children's Education Level in the United States (2021)"
 author: 
@@
 library(scales)
-#library(purrr)
-#library(forcats)
 library(patchwork)
@@
 
-In short, the findings in this paper demonstrate that parents’ education has a positive correlation with children’s education.
-Additionally, the region of living during the teenage years can also impact children’s education.
-Parents’ prestige score seems to be another factor in children’s education.
+In short, the findings in this paper demonstrate that parents' education has a positive correlation with children's education.
+Additionally, the region of living during the teenage years can also impact children's education.
+Parents' prestige score seems to be another factor in children's education.
 However, there is some limitation on the definition of prestige score that may influence the overall results.
@@
 The data will be presented clearly and succinctly with plots and tables.
-This following analysis is processed in R [@R]  with packages of tidyverse [@tidyverse], dplyr [@dplyr], here [@here], haven [@haven] and broom [@broom].
-The tables are constructed via knitr [@knitr], scales [@scales] and kableExtra [@kableExtra].
-The package inside tidyverse helps to create the plots in ggplot2 [@tidyverse].
-The missing values were imputed using the mice package [@mice].
-This paper is knitted as a PDF file by the packages of R markdown [@RMarkdown] and formatted using patchwork [@patchwork].
+This following analysis is processed in `R` [@R] with packages `tidyverse` [@tidyverse], `dplyr` [@dplyr], `here` [@here], `haven` [@haven] and `broom` [@broom].
+The tables are constructed via `knitr` [@knitr], `scales` [@scales] and kableExtra [@kableExtra].
+The package inside `tidyverse` helps to create the plots in `ggplot2` [@tidyverse].
+The missing values were imputed using `mice` [@mice].
+This paper is knitted as a PDF file by the packages of Quarto and `patchwork` [@patchwork].
 
@@
 The questionnaire includes both closed-ended and open-ended questions, but there may be concerns about biased or leading questions.
-[@norcuofch] The GSS employs various methods to adjust for non-response and ensure the representativeness of the sample.
-The limitation will be explained in detail in the @sec-weakness.
+[@norcuofch].
+The GSS employs various methods to adjust for non-response and ensure the representativeness of the sample.
+The limitations will be explained in detail in @sec-weakness.
 
 $$
-Prestige_{ij}=\mu+\alpha_i+\beta_{j}+\epsilon_{ij}
+\mbox{Prestige}_{ij}=\mu+\alpha_i+\beta_{j}+\epsilon_{ij}
 $$ {#eq-modpr}
 
-This paper will test and explore how parents’ social-economics status can affect children’s education.
+This paper will test and explore how parents' social-economics status can affect children's education.
 The data that will be used in this paper comes from the US General Social Survey from the National Opinion Research Center at the University of Chicago.
@@
 The ideal dataset would have 90,090 ratings from 1,001 raters, but due to various issues, only 82,800 ratings from 946 raters were used.
-Invalid or inconsistent ratings, such as cases with reversed codes, too few ratings, equal scores for all occupations, or a low standard deviation of ratings, were excluded to ensure the reliability of the occupational scores.[@smith]
+Invalid or inconsistent ratings, such as cases with reversed codes, too few ratings, equal scores for all occupations, or a low standard deviation of ratings, were excluded to ensure the reliability of the occupational scores [@smith].
 
-Next,the prestige score is obtained by the @eq-modpr.
-$Prestige_{ij}$ is the prestige score for occupational title i and rater j.
+Next, the prestige score is obtained by the @eq-modpr.
+$\mbox{Prestige}_{ij}$ is the prestige score for occupational title $i$ and rater $j$.
 The $\alpha_i$ represents the occupational differences of interest, while the $\beta_j$ represents differences among the raters that need to be controlled for in estimating the $\alpha_i$.
 The $\mu$ is the overall mean prestige score, and $\epsilon_{ij}$ represents the random error term.
-The range of the prestige scores is from 0 to 100, with higher scores indicating a better or more prestigious occupation[@smith].
+The range of the prestige scores is from 0 to 100, with higher scores indicating a better or more prestigious occupation [@smith].
 
@@
 
-- **mapres10**: A numeric variable indicating the respondent's mother’s prestige score.
-- **papres10**: A numeric variable indicating the respondent's father’s prestige score.
+- **mapres10**: A numeric variable indicating the respondent's mother's prestige score.
+- **papres10**: A numeric variable indicating the respondent's father's prestige score.
 - **paeduc**: A numeric variable indicating the number of years of education completed by the respondent's father.
@@
 In order to improve the representiveness of the data, this paper use the multiple-imputation and dropping methods [@gelman].
-There are two types of missing values: numercial and categorical missing values.
+There are two types of missing values: numerical and categorical missing values.
 
@@
 
-The number of oberservation before data cleaning is 4032. This number decrease to 3794 because some rows with missing values. are dropped from the data.
+The number of observations before data cleaning is 4032. This number decrease to 3794 because some rows with missing values. are dropped from the data.
 
-In addtion, the variable of **degree** is grouped into two levels.
-For respondents who have obtained at least a degree from an associate/junior college, their levels are denoted as "1", indicating they have complete a post-secondary education.
+In addition, the variable of **degree** is grouped into two levels.
+For respondents who have obtained at least a degree from an associate/junior college, their levels are denoted as "1", indicating they have completed a post-secondary education.
 For respondents who have attained only a high school diploma or possess a lower level of education, their levels are designated as "0".
@@
 #| tbl-cap: Number of respondents by degree for 2021 survey
+
 tb1 <- cleaned_data %>% count(degree) %>%
-  mutate(percent=percent(x=n/sum(n)))
-tb1<-tb1%>%arrange(factor(tb1$degree,c("graduate", "bachelor's", "associate/junior college", "high school", "less than high school")))
+  mutate(percent = percent(x = n / sum(n)))
+
+tb1 <-
+  tb1 %>% arrange(factor(
+    tb1$degree,
+    c(
+      "graduate",
+      "bachelor's",
+      "associate/junior college",
+      "high school",
+      "less than high school"
+    )
+  ))
+
 colnames(tb1) <- c("Degree", "Total", "Proportion")
-tb1%>%kable(booktabs = TRUE)
+
+tb1 %>% kable(booktabs = TRUE)
 ```
@@
 #| fig-height: 4
+
 ggplot(data = cleaned_data, aes(x = educ)) +
 
-  geom_histogram(aes(y = ..density..), binwidth = 2, fill = "white") +
+  geom_histogram(aes(y = ..density..),
+                 binwidth = 2,
+                 fill = "white") +
   geom_density(alpha = .2, fill = "green") +
@@
 ```
+
 The second variable utilized to quantify education is relation to the number of years of schooling.
-In @fig-educ, this histogram shows an overall distribution of the respondents' education in years.
+@fig-educ shows an overall distribution of the respondents' education in years.
 It is clear that the distribution is right-skewed, which indicates that there are more respondents with higher levels of education.
@@
 #| tbl-cap: Number of respondents by education for 2021 survey
-tb0 <- cleaned_data %>% summarise(Min = min(cleaned_data$educ,na.rm=T),
-                              Median = median(cleaned_data$educ,na.rm=T),
-                              Max = max(cleaned_data$educ,na.rm = T),
-                              Mean = round(mean(cleaned_data$educ,na.rm=T),digits=2))
+
+tb0 <-
+  cleaned_data %>% summarise(
+    Min = min(cleaned_data$educ, na.rm = T),
+    Median = median(cleaned_data$educ, na.rm =
+                      T),
+    Max = max(cleaned_data$educ, na.rm = T),
+    Mean = round(mean(cleaned_data$educ, na.rm =
+                        T), digits = 2)
+  )
 
@@
 ```
+
 To see a more accurate numerical summary, @tbl-educ is created.
@@
 #| fig-height: 4
-# ggplot(cleaned_data, aes(x = mapres10)) +
-#   geom_histogram(aes(y = ..density.., fill = "mapres10"), alpha = 1, binwidth = 2,color = "black") +
-#   geom_histogram(aes(x = papres10, y = ..density.., fill = "paeduc"), alpha = 0.8, binwidth = 2,color = "black") +
-#   scale_fill_manual(values = c("mapres10" = "lightblue", "paeduc" = "darkred"), name = "Variables") +
-#   labs(title = "Mother's and Father's Prestige Score",
-#        x = "Prestige Score",
-#        y = "Density")+
-#   theme_bw()
 
@@
 ggplot(cleaned_data, aes(x = mapres10)) +
-  geom_histogram(aes(y = ..density.., fill = "Mother's Prestige Score"), alpha = 1, binwidth = 1, color = "black")+
-  geom_histogram(aes(x = papres10, y = ..density.., fill = "Father's Prestige Score"), alpha = 0.8, binwidth = 1, color = "black") +
-  scale_fill_manual(values = c("Mother's Prestige Score" = "darkgreen", "Father's Prestige Score" = "lightgreen"), name = "Variables") +
+  geom_histogram(
+    aes(y = ..density.., fill = "Mother's Prestige Score"),
+    alpha = 1,
+    binwidth = 1,
+    color = "black"
+  ) +
+  geom_histogram(
+    aes(x = papres10, y = ..density.., fill = "Father's Prestige Score"),
+    alpha = 0.8,
+    binwidth = 1,
+    color = "black"
+  ) +
+  scale_fill_manual(
+    values = c(
+      "Mother's Prestige Score" = "darkgreen",
+      "Father's Prestige Score" = "lightgreen"
+    ),
+    name = "Variables"
+  ) +
   labs(x = "Prestige Score",
@@
 ```
-In @fig-parentpre, it shows the distribution of respondent’s father’s and mother’s prestige scores.
-The distribution of father’s prestige score and the mother’s prestige score have quite different distributions.
+
+@fig-parentpre shows the distribution of respondent's father's and mother's prestige scores.
+The distribution of father's prestige score and the mother's prestige score have quite different distributions.
 Specifically, the distribution of mother's prestige score is bimodal and the distribution of father's prestige score is unimodal.
@@
 #| fig-height: 4
+
 #pres vs. educ
-p1 <- ggplot(cleaned_data, aes(x=educ, y=papres10)) +
-  geom_point(color="#69b3a2", size=2,alpha=0.3) +
-  labs(x="Education (in Years)", y="Father's Prestige Score")+
+p1 <- ggplot(cleaned_data, aes(x = educ, y = papres10)) +
+  geom_point(color = "#69b3a2",
+             size = 2,
+             alpha = 0.3) +
+  labs(x = "Education (in Years)", y = "Father's Prestige Score") +
   theme_bw()
-p2 <- ggplot(cleaned_data, aes(x=educ, y=mapres10)) +
-  geom_point(color="darkgreen",size=2,alpha=0.3) +
-  labs(x="Education (in Years)", y="Mother's Prestige Score")+
+p2 <- ggplot(cleaned_data, aes(x = educ, y = mapres10)) +
+  geom_point(color = "darkgreen",
+             size = 2,
+             alpha = 0.3) +
+  labs(x = "Education (in Years)", y = "Mother's Prestige Score") +
   theme_bw()
@@
 ```
-In order to examine the relationship between parents’ prestige scores and children’s education level.
+In order to examine the relationship between parents' prestige scores and children's education level.
 The next plot, @fig-presandeduc, is used to show the relation.
-These two scatter plots show the father’s prestige score and the mother’s prestige score with their children’s education.
+These two scatter plots show the father's prestige score and the mother's prestige score with their children's education.
 It is hard to see a clear relation or pattern based on these two scatter plots.
@@
 #| fig-height: 4
+
 ggplot(cleaned_data, aes(x = maeduc)) +
-  geom_histogram(aes(y = ..density.., fill = "Mother's Education"),
-                 alpha = 1, binwidth = 0.5, color ="black") +geom_histogram(aes(x = paeduc, y = ..density.., fill = "Father's Education"),alpha = 0.8, binwidth = 0.5, color = "black") +
-scale_fill_manual(values = c("Mother's Education" = "darkgreen", "Father's Education" = "lightgreen"), name = "Variables") +
+  geom_histogram(
+    aes(y = ..density.., fill = "Mother's Education"),
+    alpha = 1,
+    binwidth = 0.5,
+    color = "black"
+  ) + geom_histogram(
+    aes(x = paeduc, y = ..density.., fill = "Father's Education"),
+    alpha = 0.8,
+    binwidth = 0.5,
+    color = "black"
+  ) +
+  scale_fill_manual(
+    values = c(
+      "Mother's Education" = "darkgreen",
+      "Father's Education" = "lightgreen"
+    ),
+    name = "Variables"
+  ) +
   labs(x = "Education (in Years)",
@@
 ```
-Another key variable is the parent’s education level.
-In @fig-parentseduc, it shows the distribution of the mother’s education and the father’s education.
+
+Another key variable is the parent's education level.
+In @fig-parentseduc, it shows the distribution of the mother's education and the father's education.
 One interesting result is that the distribution of education for mother and father is similar, or even almost identical.
@@
 ggplot(cleaned_data, aes(x = educ, fill = factor(born))) +
-  geom_histogram(aes(y = ..density..), position = "dodge", binwidth = 1.5, alpha = 0.5, color = "black")+
+  geom_histogram(
+    aes(y = ..density..),
+    position = "dodge",
+    binwidth = 1.5,
+    alpha = 0.5,
+    color = "black"
+  ) +
   labs(x = "Education (in Years)",
        y = "Density",
-       fill = "Birthplace of the Respondent (if in US)")+
+       fill = "Birthplace of the Respondent (if in US)") +
   theme_bw()
+```
 
-# ggplot(cleaned_data, aes(x = educ)) +
-# geom_histogram(aes(y = ..density.., fill = "Mother's Education"), alpha = 1, binwidth = 0.5, color = "black") +
-# geom_histogram(aes(x = paeduc, y = ..density.., fill = "Father's Education"), alpha = 0.8, binwidth = 0.5, color = "black") +
-# scale_fill_manual(values = c("Mother's Education" = "darkgreen", "Father's Education" = "lightgreen"), name = "Variables") +
-# labs(x = "Education (in Years)",
-# y = "Density") +
-# theme_bw()
-```
-Based on [@overman], the birthplace may have an impact on children’s education.
-In @fig-birth, it is used to test the relationship between birthplace and children’s education.
+@overman suggests birthplace may have an impact on children's education.
+In @fig-birth, it is used to test the relationship between birthplace and children's education.
 Note that the question of birthplace only asks if the respondent was born in the US or not.
@@
 
-```{r,echo=FALSE}
+```{r}
+#| echo: false
 tb3 <- cleaned_data %>%
   count(born) %>%
-  mutate(percent=percent(x=n/sum(n)))
+  mutate(percent = percent(x = n / sum(n)))
 ```
@@
 #| tbl-cap: The proportion and mean of education for different regions
+
 tb_reg <- cleaned_data %>%
   count(reg16) %>%
-  mutate(percent=percent(x=n/sum(n)))
+  mutate(percent = percent(x = n / sum(n)))
 
-tb_reg_avg<-cleaned_data %>%
+tb_reg_avg <- cleaned_data %>%
   group_by(reg16) %>%
-  summarise(mean_educ = round(mean(educ),digits=2))
-tb_reg<-tb_reg%>%mutate("Mean of Education"=tb_reg_avg$mean_educ)
-colnames(tb_reg)<-c("Region","Count","Percent","Mean of Education")
+  summarise(mean_educ = round(mean(educ), digits = 2))
+tb_reg <- tb_reg %>% mutate("Mean of Education" = tb_reg_avg$mean_educ)
+colnames(tb_reg) <- c("Region", "Count", "Percent", "Mean of Education")
 tb_reg %>% kable()
 ```
-The @tb-regandeduc showcases the average number of years of education for individuals in different living regions of the United States when the respondent's was 16.
+
+@tb-regandeduc showcases the average number of years of education for individuals in different living regions of the United States when the respondent was 16.
 The highest average education level is found in the Middle Atlantic region, where the mean is 15.13 years.
@@
 #| fig-height: 4
-ggplot(cleaned_data, aes(x = educ,fill = factor(reg16))) +
-  geom_histogram(aes(y = ..density..),binwidth = 2, color = "black") +
-  facet_wrap(~ reg16, ncol = 3)+
+
+ggplot(cleaned_data, aes(x = educ, fill = factor(reg16))) +
+  geom_histogram(aes(y = ..density..),
+                 binwidth = 2,
+                 color = "black") +
+  facet_wrap( ~ reg16, ncol = 3) +
   labs(x = "Education (in Years)",
-       y = "Density")+
-  theme_bw()+
+       y = "Density") +
+  theme_bw() +
   theme(legend.position = "none")
 ```
-According to the study [@nieuwenhuis], there appears to be an association between neighborhood and education.
+
+@nieuwenhuis finds an association between neighborhood and education.
 In the 2021 GSS, the neighbourhood variable is being measured by the living region at the age of 16.
@@
 #| fig-height: 4
+
 ggplot(cleaned_data, aes(x = educ, y = maeduc, color = incom16)) +
@@
   stat_smooth(method = "lm", se = FALSE) +
-  facet_wrap(~incom16)+
+  facet_wrap( ~ incom16) +
   labs(x = "Education (in Years)",
@@
 To examine the relationship between family income and education, the @fig-maincome and @fig-paincome visualize this relation.
-In @fig-maincome, the vertical axis represents the mother’s education and the horizontal axis represents the children’s education.
+In @fig-maincome, the vertical axis represents the mother's education and the horizontal axis represents the children's education.
 The respondents were asked to rate their family income at age of 16 in the following category: (1) **far above average**, (2) **above average**, (3) **average**, (4) **below average** and (5) **far below average**.
-Overall, mothers’ education and children’s education are positively correlated despite the level of income.
+Overall, mothers' education and children's education are positively correlated despite the level of income.
 To be more specific, for respondents whose family income was **far above average**, the slope seems to be steeper than the other groups, which indicates that for those with higher family income, there may be a stronger relationship between the education of the mother and the education of the children.
@@
 ```
+
 In @fig-paincome, the income groups are the same as before.
-However, the plot is now examining the correlation between the father’s education and the children’s education.
+However, the plot is now examining the correlation between the father's education and the children's education.
 Once again, there is a positive correlation between the education level of fathers and their children's education, regardless of family income.
@@
 
-# Model* {#sec-model}
+# Model {#sec-model}
+
+We estimate the following mode:
 
 $$
-\log(\frac{\pi}{1-\pi})=\hat{\beta_0}+\hat{\beta_1}X_{mapres10}+\hat{\beta_2}X_{papres10}+\hat{\beta_3}X_{paeduc}+\hat{\beta_4}X_{maeduc}
+\log\left(\frac{\pi}{1-\pi}\right)=\hat{\beta_0}+\hat{\beta_1}X_{\mbox{mapres10}}+\hat{\beta_2}X_{\mbox{papres10}} + \hat{\beta_3}X_{\mbox{paeduc}} + \hat{\beta_4}X_{\mbox{maeduc}}
 $$ {#eq-model}
@@
 # Fit the model
-model <- glm(high_school ~ mapres10 + papres10 + paeduc + maeduc + incom16 + reg16, family="binomial", data = cleaned_data)
+model <-
+  glm(
+    high_school ~ mapres10 + papres10 + paeduc + maeduc + incom16 + reg16,
+    family = "binomial",
+    data = cleaned_data
+  )
 
-modelsummary::modelsummary(
-  list(
-    "Children's Education" = model
-  ),
+modelsummary::modelsummary(list("Children's Education" = model),
                            fmt = 2,
-  statistic = "conf.int"
-)
+                           statistic = "conf.int")
 ```
+
 The @tbl-lmeduc is built based on the @eq-model. In this logistic model, it is used to predict the probability of a child attains a post-secondary degree based on these predictors.
@@
 
-From @tbl-lmeduc,the positive coefficients for both mother's and father's prestige scores (**mapres10** and **papres10**) suggest a slight positive relationship with children's education levels.
+From @tbl-lmeduc, the positive coefficients for both mother's and father's prestige scores (**mapres10** and **papres10**) suggest a slight positive relationship with children's education levels.
 Moreover, the positive coefficients for parents' education (**paeduc** and **maeduc**) have a greater positive association with children's education level.
@@
 
-# Discussion*
+# Discussion
 
 ## Average years of schooling in the US and Outside of US
+
 Education has been a crucial driver of social and economic progress worldwide, and its expansion over the past two centuries has been a significant phenomenon.
 The global average year of schooling currently stands at around 8.7 years, according to data from World Economics [@world_economics].
+
 In this paper, by examining the average year of schooling in the United States, based on 2021 GSS data, it was found that the average year of schooling is 14 in @tbl-educ.
@@
 ## The impact of parental education level
-In the @sec-result, the @fig-parentseduc, @fig-paincome and @fig-maincome, they all show that there is a postive correlation between the education levels of mothers and fathers.
+
+In the @sec-result, the @fig-parentseduc, @fig-paincome and @fig-maincome, they all show that there is a positive correlation between the education levels of mothers and fathers.
 If one parent has a high level of education, it is likely that the other parent also has a high level of education.
@@
 Studies have consistently shown that parents' education is one of the strongest predictors of their children's academic achievement and educational attainment [@hout].
-The study by  [@chetty] used data from tax records and college enrollment records to analyse the relationship between parents' income, education, and social status, and their children's earnings and educational outcomes.
+@chetty used data from tax records and college enrollment records to analyse the relationship between parents' income, education, and social status, and their children's earnings and educational outcomes.
 The results showed that children from high-income families were more likely to attend college and earn higher incomes than children from low-income families.
@@
 Furthermore, the analysis of the statistical model from the @tbl-lmeduc shows that parental education and prestige scores are positively correlated with their children's academic achievement and societal status.
-The coefficients for the independent variables father’s and mother’s education in years indicate that an increase in fathers' education has a greater association with their children's education than that of mothers.
+The coefficients for the independent variables father's and mother's education in years indicate that an increase in fathers' education has a greater association with their children's education than that of mothers.
 This suggests that the father's role in education and investment in their children's future may be particularly important.
@@
 Despite these discrepancies, the findings in this paper concur with prior literature that emphasizes the importance of parental education levels and their impact on college student achievement [@ortiz].
-This study focuses on showing that the fathers' education level has a significant relationship with childrens’ year of schooling, while no statistically significant relationship was observed with the mothers' education level.
+This study focuses on showing that the fathers' education level has a significant relationship with childrens' year of schooling, while no statistically significant relationship was observed with the mothers' education level.
 
@@
 
-### Weaknesses and Limitations
+### Weaknesses
+
 As with any research, this study has its limitations and weaknesses.
@@
 ### Next Steps
+
 To shed light in future studies, there are some potential improvements.

### 2023-04-17 outputs/paper/paper.qmd
final draft of the paper

@@
 
-In the past, family status variables, such as the socio-economic status and educational level of parents, were commonly considered as strong predictors of academic achievement in children.
-More specifically, parents who have lower levels of education, and those with higher levels of education are more inclined to view higher education as desirable and encourage their children to excel academically.
-They also tend to hold higher expectations for their children's academic performance [@davis_2014].
-Ultimately, these may affect the education level of the children.
+Education plays a crucial role in shaping an individual's life chances, personal development, and socioeconomic mobility.
+Over the years, researchers have sought to understand the factors that may influence children's academic achievement and education levels, with a particular focus on the impact of family background variables such as parents' socio-economic status and educational attainment [@coleman; @sirin].
+To reveal the complex relationship between family background and children's education, it is important to developing effective educational policies and interventions that promote equal opportunities for all students, regardless of their social or economic origins.
 
-One potential explanation for this relationship is that higher levels of education can provide parents with access to resources such as income, time, energy, and community contacts, which can enable greater involvement in a child's education.
-As a result, the influence of family status variables on children's academic achievement may be best understood as a complex interaction between status and process variables [@khan].
+Several studies have documented the strong and consistent association between parents' education and children's academic outcomes [@haveman].
+Parents with higher education levels are more likely to provide an intellectually stimulating home environment, hold higher expectations for their children's academic performance, and be actively involved in their education, thus fostering their children's academic success [@davis2].
+However, recent research has highlighted the multifaceted nature of this relationship, indicating that it can be influenced by various factors such as family income, social class, and so on [@reardon, @domina].
 
-However, recent research has indicated that the relationship between these factors and academic achievement is not always direct.
-Instead, socio-economic status and parents' education are part of a larger set of psychological and sociological variables that can impact children's educational outcomes [@khan].
+Despite the substantial body of research on this topic, there is still much to learn about the intricate ways in which family background affects children's education levels, particularly in the context of growing economic inequality and social stratification [@chetty2].
+This paper aims to contribute to the ongoing scholarly discourse by examining the relationship between family status variables, including parents' socio-economic status and educational attainment, and their impact on children's academic achievement and education level using data from the 2021 General Social Survey (GSS).
+By gaining a deeper understanding of these dynamics,policymakers will be better equipped to address educational disparities and work toward fostering more equitable opportunities for all students.
 
@@
 
-@tb-regandeduc showcases the average number of years of education for individuals in different living regions of the United States when the respondent was 16.
+@tbl-regandeduc showcases the average number of years of education for individuals in different living regions of the United States when the respondent was 16.
 The highest average education level is found in the Middle Atlantic region, where the mean is 15.13 years.
@@
 In the Pacific and New England regions, the mean years of education are relatively similar, with 14.94 and 14.85 years, respectively.
-Those regions are above the overall average, which is 14.82 from @tb_educ.
+Those regions are above the overall average, which is 14.82 from @tbl-educ.
 On the other hand, the South Atlantic and West North Central regions also display comparable average education levels of 14.75 and 14.73 years, respectively.
@@
 ```
+$$
+\mbox{Odds}=\frac{\pi}{1-\pi}=\mbox{exp}(\hat{\beta_0}+\hat{\beta_1}X_{\mbox{mapres10}}+\hat{\beta_2}X_{\mbox{papres10}} + \hat{\beta_3}X_{\mbox{paeduc}} + \hat{\beta_4}X_{\mbox{maeduc}})
+$$ {#eq-coef}
 
 The @tbl-lmeduc is built based on the @eq-model. In this logistic model, it is used to predict the probability of a child attains a post-secondary degree based on these predictors.
-*missing the bullet point to explain each exp(coef)*
 
-From @tbl-lmeduc, the positive coefficients for both mother's and father's prestige scores (**mapres10** and **papres10**) suggest a slight positive relationship with children's education levels.
-Moreover, the positive coefficients for parents' education (**paeduc** and **maeduc**) have a greater positive association with children's education level.
+Using the @eq-coef, the coefficients of the model can be interpreted as follow and note that the base group for income level is `Average` and for region is `East North Central`:
 
-The coefficients for the income level at age of 16 (**incom16**) indicate that higher-income levels are generally associated with lower education levels for children.
-Particularly, the "far above average" and "far below average" income categories show a significant negative relationship with children's education levels.
+- **Intercept**: The odds of a child completing high school when all other predictor variables are at their reference levels (i.e., income level is 'average' and region is 'east north') is exp(-2.59) = 0.075, with a 95% confidence interval of [exp(-2.99), exp(-2.19)] = [0.05, 0.11].
 
-Regarding the living regional differences at age of 16 (**reg16**), the results indicate that children from the east south atlantic region may have lower education levels compared to those from other regions.
-The positive coefficients for middle atlantic and west north central regions suggest a positive association with children's education levels, although the results are not statistically significant.
+- **Respondent's Mother's Prestige Score (mapres10)**: A one-unit increase in the respondent's mother's prestige score is associated with an increase in the odds of a child completing high school by exp(0.01) = 1.01, with a 95% confidence interval of [exp(0.01), exp(0.02)] = [1.01, 1.02].
+
+- **Respondent's Father's Prestige Score (papres10)**: A one-unit increase in the respondent's father's prestige score is associated with an increase in the odds of a child completing high school by exp(0.02) = 1.02, with a 95% confidence interval of [exp(0.01), exp(0.02)] = [1.01, 1.02].
+
+- **Respondent's Father's Education (paeduc)**: A one-year increase in the number of years of education completed by the respondent's father is associated with an increase in the odds of a child completing high school by exp(0.10) = 1.11, with a 95% confidence interval of [exp(0.07), exp(0.12)] = [1.07, 1.13].
+
+- **Respondent's Mother's Education (maeduc)**: A one-year increase in the number of years of education completed by the respondent's mother is associated with an increase in the odds of a child completing high school by exp(0.03) = 1.03, with a 95% confidence interval of [exp(0.01), exp(0.06)] = [1.01, 1.06].
+
+- **Respondent's Income Level above Average (incom16above average)**: The odds of a child completing high school for respondents with an above average household income at age 16 are exp(0.29) = 1.34 times the odds for respondents with an average household income, with a 95% confidence interval of [exp(0.09), exp(0.50)] = [1.09, 1.65].
+
+- **Respondent's Income Level below Average (incom16below average)**: The odds of a child completing high school for respondents with a below-average household income at age 16 are exp(-0.11) = 0.90 times the odds for respondents with an average household income, with a 95% confidence interval of [exp(-0.28), exp(0.06)] = [0.75, 1.06].
+
+- **Respondent's Income Level Far above Average (incom16far above average)**: The odds of a child completing high school for respondents with a far above-average household income at age 16 are exp(0.28) = 1.33 times the odds for respondents with an average household income, with a 95% confidence interval of [exp(-0.20), exp(0.78)] = [0.82, 2.18].
+
+- **Respondent's Income Level Far below Average (incom16far below average)**: The odds of a child completing high school for respondents with a far below-average household income at age 16 are exp(-0.21) = 0.81 times the odds for respondents with an average household income, with a 95% confidence interval of [exp(-0.44), exp(0.03)] = [0.64, 1.03].
+
+- **Respondent's Living Region in East South Atlantic at 16 (reg16east south atlantic)**: The odds of a child completing high school for respondents living in the East South Atlantic region at age 16 are exp(-0.18) = 0.83 times the odds for respondents living in the East North region, with a 95% confidence interval of [exp(-0.49), exp(0.14)] = [0.61, 1.15].
+
+- **Respondent's Living Region in Middle Atlantic at 16 (reg16middle atlantic)**: The odds of a child completing high school for respondents living in the Middle Atlantic region at age 16 are exp(0.26) = 1.30 times the odds for respondents living in the East North region, with a 95% confidence interval of [exp(0.02), exp(0.50)] = [1.02, 1.65].
+
+- **Respondent's Living Region in Mountain at 16 (reg16mountain)**: The odds of a child completing high school for respondents living in the Mountain region at age 16 are exp(-0.19) = 0.83 times the odds for respondents living in the East North region, with a 95% confidence interval of [exp(-0.51), exp(0.12)] = [0.60, 1.13].
+
+- **Respondent's Living Region in New England at 16 (reg16new england)**: The odds of a child completing high school for respondents living in the New England region at age 16 are exp(0.09) = 1.09 times the odds for respondents living in the East North region, with a 95% confidence interval of [exp(-0.26), exp(0.45)] = [0.77, 1.57].
+
+- **Respondent's Living Region in Pacific at 16 (reg16pacific)**: The odds of a child completing high school for respondents living in the Pacific region at age 16 are exp(0.02) = 1.02 times the odds for respondents living in the East North region, with a 95% confidence interval of [exp(-0.20), exp(0.24)] = [0.82, 1.27].
+
+- **Respondent's Living Region in South Atlantic at 16 (reg16south atlantic)**: The odds of a child completing high school for respondents living in the South Atlantic region at age 16 are exp(-0.06) = 0.94 times the odds for respondents living in the East North region, with a 95% confidence interval of [exp(-0.30), exp(0.18)] = [0.74, 1.20].
+
+- **Respondent's Living Region in West North Central at 16 (reg16west north central)**: The odds of a child completing high school for respondents living in the West North Central region at age 16 are exp(-0.16) = 0.85 times the odds for respondents living in the East North region, with a 95% confidence interval of [exp(-0.43), exp(0.12)] = [0.65, 1.13].
+
+- **Respondent's Living Region in West South Central at 16(reg16west south central)**: The odds of a child completing high school for respondents living in the West South Central region at age 16 are exp(-0.33) = 0
 
@@
 
+## The impact of family income level and prestige score
+
+From @tbl-lmeduc, it show the coefficients for the income level at age of 16 and indicates that higher-income levels are generally associated with lower education levels for children.
+Particularly, the "far below average" income categories show a significant negative relationship with children's education levels in terms of the odds of attaining a post-secondary degree.
+
+The correlation between parents' education levels and their children's educational outcomes is well-established in the literature.
+Studies have consistently shown that parents' education is one of the strongest predictors of their children's academic achievement and educational attainment [@hout].
+@chetty used data from tax records and college enrollment records to analyse the relationship between parents' income, education, and social status, and their children's earnings and educational outcomes.
+The results showed that children from high income families were more likely to attend college and earn higher incomes than children from low income families.
+The findings from the paper match with the existing research, that higher income level has a postive effect on the children's education level.
+
+However, the coefficients for parents' (both mother's and father's) prestige score show a neglectable positive effect.
+The effect is too small to draw an remarkable conclusion.
+Yet, the purpose of using the prestige score is to measure the family socialeconomic status based on their occupation.
+There are some reason to explain why the effects are not significant and @sec-weakness will provide some potential solution to better examine this relationship.
+
 ## The impact of parental education level
@@
 
-The correlation between parents' education levels and their children's educational outcomes is well-established in the literature.
-Studies have consistently shown that parents' education is one of the strongest predictors of their children's academic achievement and educational attainment [@hout].
-@chetty used data from tax records and college enrollment records to analyse the relationship between parents' income, education, and social status, and their children's earnings and educational outcomes.
-The results showed that children from high-income families were more likely to attend college and earn higher incomes than children from low-income families.
-In addition, the study found that parental education was a key factor in determining children's educational and economic outcomes.
-Children with highly educated parents were more likely to attend college and earn higher incomes than children with less educated parents.
-
-Furthermore, the analysis of the statistical model from the @tbl-lmeduc shows that parental education and prestige scores are positively correlated with their children's academic achievement and societal status.
+Furthermore, the analysis of the logistic model from the @tbl-lmeduc shows that parental education and prestige scores are positively correlated with their children's probability of attaining a post-secondary degree.
 The coefficients for the independent variables father's and mother's education in years indicate that an increase in fathers' education has a greater association with their children's education than that of mothers.
@@
 However, the influence of both parents is still significant.
-The R-squared value suggests that the independent variables explain only a small portion of the variation in the dependent variable, indicating that other factors beyond parental education and prestige score also play a role.
-Nonetheless, these findings highlight the importance of parental involvement in their children's education, with a particular emphasis on the role of fathers.
+Nonetheless, these findings highlight the importance of parental education level in their children's education, with a particular emphasis on the role of fathers.
 
@@
 The present study showed that paternal support of education has a significant impact on academic achievement, while maternal education level did not show a significant impact.
-However, the study of @2000 showed that families with active fathers fostered maternal involvement in the household, which resulted in a support system that fosters educational attainment and positive long-term contributions.
+Additionally, the study of @2000 showed that families with active fathers fostered maternal involvement in the household, which resulted in a support system that fosters educational attainment and positive long-term contributions.
+This suggest that the role of parents both contribute to children's education.
 
-Despite these discrepancies, the findings in this paper concur with prior literature that emphasizes the importance of parental education levels and their impact on college student achievement [@ortiz].
-This study focuses on showing that the fathers' education level has a significant relationship with childrens' year of schooling, while no statistically significant relationship was observed with the mothers' education level.
+The findings in this paper concur with prior literature that emphasizes the importance of parental education levels and their impact on college student achievement [@ortiz].
+This study focuses on showing that the fathers' education level has a significant relationship with childrens' year of schooling, while no  significant relationship was observed with the mothers' education level.
+
+In addition, the study found that parental education was a key factor in determining children's educational and economic outcomes.
+Children with highly educated parents were more likely to attend college and earn higher incomes than children with less educated parents [@chetty].
+
+## The Effect of Living Region during the Adolescent Period
+
+Regarding the living regional differences at age of 16 from @tbl-lmeduc, the coefficients indicate that children from the East South Atlantic, West South Central, Mountain and West North Central may have a negative effect on completed a post-secondary degree compared to those from other regions.
+The positive coefficients for Middle Atlantic regions and New England suggest a positive association with children's education levels, although the results are not statistically significant.
+The @eq-model employs the probability of a child obtaining a post-secondary degree as the response variable.
+A potential explanation for the positive coefficient could be that certain regions offer higher-quality education, particularly in the form of high schools.
+
+According to the US News rankings in 2022, these are the top 8 high schools in the US.
+Thomas Jefferson High School for Science and Technology is located in the South Atlantic region in Virginia, as is the Academic Magnet High School in South Carolina and the School for Advanced Studies (SAS) in Florida.
+The Signature School can be found in the East North Central region in Indiana, while the Payton College Preparatory High School is also located in the East North Central region in Illinois.
+The Davidson Academy of Nevada is situated in the Mountain region in Nevada.
+The Central Magnet School is in the East South Central region in Tennessee, and The School for the Talented and Gifted (TAG) is located in the West South Central region in Texas.
+[@best]
+
+It's noteworthy that none of the top schools are from the Middle Atlantic or New England regions, which have positive coefficients in the model.
+This discrepancy could be due to differences in time.
+School rankings can change over the years as states develop, and the GSS respondents were born in a different time period.
+This leads to a mismatch between the rankings and the model's coefficients.
 
@@
 Therefore, further research is necessary to fully understand the complex relationship between these variables and their impact on academic achievement, which will be carefully discussed in @sec-weakness.
-
-## The Effect of Living Region during the Adolescent Period
-referring results from the model about how regions affect the education level
 

### 2023-04-17 scripts/02-data_cleaning.R

@@ -75,6 +75,9 @@ cleaned_data <- cleaned_data %>%
 cleaned_data <- cleaned_data %>%
   mutate(high_school = ifelse(degree %in% c("graduate", "bachelor's", "associate/junior college"),1,0))
 
+cleaned_data$incom16 <- as.factor(cleaned_data$incom16)
+cleaned_data$incom16 <- relevel(cleaned_data$incom16, ref = "average")
+
 #### Save data ####
 write_csv(cleaned_data, "inputs/data/cleaned_GSS2021.csv")
 

### 2023-04-17 outputs/paper/paper.qmd
tiny changes in the data cleaning

@@
 For respondents who have obtained at least a degree from an associate/junior college, their levels are denoted as "1", indicating they have completed a post-secondary education.
-For respondents who have attained only a high school diploma or possess a lower level of education, their levels are designated as "0".
+This is because the logistic model only takes binary response variable.
+Hence, for respondents who have attained only a high school diploma or possess a lower level of education, their levels are labelled as "0".
 

### 2023-04-18 outputs/paper/paper.qmd
small changes

@@
 
-# Data {#sec-data}
+417 add a paragraph for each section
+
+# Data {#sec-data} (417 add the graph for data)
 
@@
 
-The number of observations before data cleaning is 4032. This number decrease to 3794 because some rows with missing values. are dropped from the data.
+The number of observations before data cleaning is 4032. This number decreases to 3794 because some rows with missing values. are dropped from the data.
 
@@
 #| label: fig-educ
-#| fig-cap: Histogram of Respondents' Education
+#| fig-cap: Distribution of Respondents' Education
 #| echo: false
@@
 tb_reg <- tb_reg %>% mutate("Mean of Education" = tb_reg_avg$mean_educ)
-colnames(tb_reg) <- c("Region", "Count", "Percent", "Mean of Education")
+colnames(tb_reg) <- c("Region", "Count", "Proportion", "Mean of Education")
 tb_reg %>% kable()
@@
 #| label: fig-region
-#| fig-cap: Respondent's Living Region at Age of 16 v.s. Education
+#| fig-cap: Respondent's Living Region at Age of 16 with Education
 #| echo: false
@@
 The Davidson Academy of Nevada is situated in the Mountain region in Nevada.
-The Central Magnet School is in the East South Central region in Tennessee, and The School for the Talented and Gifted (TAG) is located in the West South Central region in Texas.
-[@best]
+The Central Magnet School is in the East South Central region in Tennessee, and The School for the Talented and Gifted (TAG) is located in the West South Central region in Texas [@best].
 
@@
 This study opted not to use the linear model due to violations of underlying assumptions.
-Future research may improve upon this by constructing a more sophisticated linear regression model to elucidate the relationship between these variables in a more precise manner.
+Future research may improve upon this by constructing a more sophisticated linear regression model to demonstrate the relationship between these variables in a more precise manner.
 

### 2023-04-18 outputs/paper/paper.qmd
changes based on the comments

@@
 
-417 add a paragraph for each section
+The structure of this paper comprises several sections.
+In the @sec-data, it introduces the source of the data and the cleaning process, as well as the characteristics of the variables.
+In the @sec-result, it reveals some findings from the data regarding how parents' backgrounds may affect children's education levels.
+In the @sec-discussion, it take a closer look at the results and compares them with the published researches.
+Additionally, the @sec-weakness lists some of the potential limitations of the paper and sheds light on future studies.
+The section of @sec-conclusion is the conclusion of the paper.
+Lastly, a mock survey related to the topic of this paper can be found in the @sec-appendix.
 
-# Data {#sec-data} (417 add the graph for data)
+# Data {#sec-data}
 
 ## Source
+
 The 2021 General Social Survey (GSS) is a nationally representative survey conducted to collect data on social trends and attitudes among people living in the United States.
@@
 
-# Results {#sec-result}
-
-This following findings are from the 2021 General Social Survey (GSS) data.
+## Characteristics of the variables
 
@@
 ```{r}
-#| label: fig-presandeduc
-#| fig-cap: Mother's and Father's Prestige Score with children's education
-#| echo: false
-#| warning: false
-#| message: false
-#| fig-width: 8
-#| fig-height: 4
-
-#pres vs. educ
-p1 <- ggplot(cleaned_data, aes(x = educ, y = papres10)) +
-  geom_point(color = "#69b3a2",
-             size = 2,
-             alpha = 0.3) +
-  labs(x = "Education (in Years)", y = "Father's Prestige Score") +
-  theme_bw()
-p2 <- ggplot(cleaned_data, aes(x = educ, y = mapres10)) +
-  geom_point(color = "darkgreen",
-             size = 2,
-             alpha = 0.3) +
-  labs(x = "Education (in Years)", y = "Mother's Prestige Score") +
-  theme_bw()
-
-p1 + p2
-```
-In order to examine the relationship between parents' prestige scores and children's education level.
-The next plot, @fig-presandeduc, is used to show the relation.
-These two scatter plots show the father's prestige score and the mother's prestige score with their children's education.
-It is hard to see a clear relation or pattern based on these two scatter plots.
-To verify if the prestige score is one of the factors, a linear model will be used in the @sec-model.
-
-```{r}
 #| label: fig-parentseduc
@@
 
+# Results {#sec-result}
+
+This following findings are from the 2021 General Social Survey (GSS) data.
+
+```{r}
+#| label: fig-presandeduc
+#| fig-cap: Mother's and Father's Prestige Score with children's education
+#| echo: false
+#| warning: false
+#| message: false
+#| fig-width: 8
+#| fig-height: 4
+
+#pres vs. educ
+p1 <- ggplot(cleaned_data, aes(x = educ, y = papres10)) +
+  geom_point(color = "#69b3a2",
+             size = 2,
+             alpha = 0.3) +
+  labs(x = "Education (in Years)", y = "Father's Prestige Score") +
+  theme_bw()
+p2 <- ggplot(cleaned_data, aes(x = educ, y = mapres10)) +
+  geom_point(color = "darkgreen",
+             size = 2,
+             alpha = 0.3) +
+  labs(x = "Education (in Years)", y = "Mother's Prestige Score") +
+  theme_bw()
+
+p1 + p2
+```
+
+In order to examine the relationship between parents' prestige scores and children's education level.
+The next plot, @fig-presandeduc, is used to show the relation.
+These two scatter plots show the father's prestige score and the mother's prestige score with their children's education.
+It is hard to see a clear relation or pattern based on these two scatter plots.
+To verify if the prestige score is one of the factors, a linear model will be used in the @sec-model.
+
 ```{r}
@@
 
-# Model {#sec-model}
+## Model {#sec-model}
 
@@
 
-# Discussion
+# Discussion {#sec-discussion}
 
@@
 
-In this paper, by examining the average year of schooling in the United States, based on 2021 GSS data, it was found that the average year of schooling is 14 in @tbl-educ.
-This result is consistent with the findings from Our World in Data, which indicate that the average year of schooling in the US is also 14 years [@global_education].
+In the previous paper, the average year of schooling in the United States, based on 2021 GSS data, it was found that the average year of schooling is 14.76 and the avergae now is 14.82 in @tbl-educ.
+Despite the tiny changes in numbers due to the multiple imputation, this result is still consistent with the findings from Our World in Data, which indicate that the average year of schooling in the US is also 14 years [@global_education].
 
@@
 
-It's noteworthy that none of the top schools are from the Middle Atlantic or New England regions, which have positive coefficients in the model.
+It is noteworthy that none of the top schools are from the Middle Atlantic or New England regions, which have positive coefficients in the model.
 This discrepancy could be due to differences in time.
@@
 
-## Weaknesses and next steps {#sec-weakness}
+# Weaknesses and next steps {#sec-weakness}
 
-### Weaknesses
+## Weaknesses
 
@@
 
-### Next Steps
+## Next Steps
 
@@
 
-# Appendix
+# Conclusion {#sec-conclusion}
+
+Education is a critical component in both social and economic progress.
+Therefore, understanding educational trends in the US has valuable implications for policymaking and decision-making.
+This paper examines the impact of family income level, parental education, and living regions during adolescence on children's education.
+Higher income levels are generally associated with higher education levels for children, and fathers' education levels have a stronger impact on their children's academic achievement than mothers' education levels.
+The influence of both parents is still significant.
+
+Adolescent living environments also influence completion of a postsecondary degree.
+Although the results are not statistically significant, the Middle Atlantic and New England regions demonstrate a favourable connection with children's educational levels.
+These findings highlight the significance of parental involvement and support in children's education, with a particular emphasis on the role of fathers.
+Future research is needed to properly comprehend the complex relationship between these variables and their impact on children's educational attainment.
+
+# Appendix {#sec-appendix}
 
